\section{Numerical Verification and Audit}
\label{sec:numerical-verification}

This chapter orders the available numerical evidence for the implemented solver.
The evidence supports bounded implementation-verification claims for the declared
operator and recorded diagnostics. It does not support validation, accuracy,
pressure-drop, FFR, physiological, clinical, predictive, or causal claims.

\subsection{Evidence Hierarchy and Scope}
\label{subsec:verification-scope-convergence-stability}

The hierarchy used below is explicit. The strongest positive evidence is the
manufactured-solution record, because it compares the implemented forced
operator with a declared smooth exact solution. The next and principal numerical
result is the geometry-rest preservation test, because the selected comparison
runs start from the same rest state and the current MUSCL/Rusanov realization
fails to preserve it. Boundary, CFL, positivity, and conservation diagnostics
then state whether accepted runs remained within the recorded numerical regime.
Self-convergence, backend parity, stationary-Stokes projection, broad
rheology/profile sweeps, resolved-velocity benchmark stages, and nonselected DG
rows are secondary implementation-health records and are kept in
Appendix~\ref{app:num-secondary-implementation-health}.

Standard finite-volume context is supplied by conservative hyperbolic
discretization with consistent numerical fluxes and sources, CFL-controlled time
stepping, limiter and Rusanov dissipation choices, and SSP Runge--Kutta time
integration \cite{LeVeque2002FiniteVolumeMethodsForHyperbolicProblems}. In 1D
blood-flow practice, MUSCL/Rusanov and Lax--Friedrichs-type discretizations are
commonly assessed through exact-solution, benchmark, grid-refinement, and
equilibrium tests
\cite{Wang2014BloodFlowNetworks,BeckersKolbe2025LaxFriedrichsHemodynamics,BoileauEtAl2015Benchmark,BenemeritoEtAl2024OpenBF,LuccaEtAl2025FFRPrediction}.
Here those records are read as empirical implementation evidence, not as a
formal convergence theorem for the full nonlinear stenosis solver.

\subsection{Manufactured-Solution Evidence}
\label{subsec:mms-verification}

The implementation exposes an opt-in manufactured forcing term. The production
equations use \texttt{NoForcing()} by default. For verification only,
\texttt{ManufacturedForcing()} supplies a smooth exact state
$(a^\star(z,t),q^\star(z,t))$, exact initial data, exact boundary states, and
source terms inserted into the existing finite-volume right-hand side. In solver
coordinates, let
\begin{flalign*}
    \quad
    s(z)&=\sin\!\left(\frac{\pi z}{L}\right),&
    a_0(z)&=R_0(z)^2,&
    \omega&=\frac{2\pi}{T_{\mathrm{mms}}}. &&
\end{flalign*}
The default manufactured record uses
\begin{flalign*}
    \quad
    \epsilon_a&=0.02,&
    U_{\mathrm{mms}}&=2.0\,\mathrm{cm}/\mathrm{s},&
    T_{\mathrm{mms}}&=0.02\,\mathrm{s}, &&
\end{flalign*}
and defines
\begin{flalign*}
    \quad
    a^\star(z,t)
    &=
    a_0(z)\left[1+\epsilon_a s(z)\cos(\omega t)\right], &&\\
    \quad
    q^\star(z,t)
    &=
    a_0(z)U_{\mathrm{mms}}s(z)^2\sin(\omega t). &&
\end{flalign*}
The tabulated run uses $0\le z\le L=6\,\mathrm{cm}$ and
$0\le t\le2\times10^{-4}\,\mathrm{s}$ unless the CLI options below are changed.
Initial data are sampled from $(a^\star,q^\star)$ at $t=0$, and the boundary
state routine receives the exact traces at $z=0$ and $z=L$. The inserted forcing
terms are the residuals required by the recorded flux and source realization,
\begin{flalign*}
    \quad
    f_a(z,t)
    &=
    \pd_t a^\star+\pd_z F_a(a^\star,q^\star;z), &&\\
    \quad
    f_q(z,t)
    &=
    \pd_t q^\star+
    \pd_z F_q(a^\star,q^\star;z)-S_q(a^\star,q^\star;z), &&
\end{flalign*}
where $F_a,F_q$ and $S_q$ are the selected discrete-model point flux and source
functions. The spatial derivatives used to evaluate the forcing are generated by
the verification helper, so the test checks the implemented operator against
this declared manufactured record rather than a physical stenosis experiment.

\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
bin/stenosis-hemodynamics verify mms \
  --nxs 50,100,200,400 \
  --dt-values 2e-5,1e-5,5e-6,2.5e-6 \
  --output-dir report/assets/tables/verification \
  --overwrite
\end{lstlisting}

The runner records cell-center discrete errors
\begin{flalign*}
    \quad
    \|e\|_1&=\frac{1}{N}\sum_i |e_i|,&
    \|e\|_2&=\left(\frac{1}{N}\sum_i e_i^2\right)^{1/2},&
    \|e\|_\infty&=\max_i |e_i|, &&
\end{flalign*}
with observed orders computed from the $L^2$ rows as
\begin{flalign*}
    \quad
    p_{\mathrm{obs}}
    &=
    \frac{\log(E_{\mathrm{coarse}}/E_{\mathrm{fine}})}
         {\log(\Delta_{\mathrm{coarse}}/\Delta_{\mathrm{fine}})}. &&
\end{flalign*}
For spatial rows $\Delta$ is the grid spacing. For requested-timestep rows the
native solver also records realized CFL diagnostics because the accepted
timestep is $\min(\Delta t_{\mathrm{request}},\Delta t_{\mathrm{CFL}})$. The
temporal rows in the generated MMS table are therefore read as a
time-step-insensitivity check at the displayed spatial resolution, not as a
clean temporal-order study.

\IfFileExists{report/assets/tables/verification/mms_verification.tex}{%
    \input{report/assets/tables/verification/mms_verification.tex}
}{%
    \begin{center}
    \small
    \emph{Manufactured-solution table unavailable in this checkout. Run
    \texttt{verify mms} to populate
    \path{report/assets/tables/verification/}.}
    \end{center}
}
The spatial rows provide bounded, positive implementation-verification evidence
for the declared forced operator. The observed rates are below the formal
smooth-region MUSCL/SSPRK3 expectation, so this table is not reported as a
formal-order result for every model variant. The $L^1$ history records
integrated error behavior, while $L^\infty$ highlights localized boundary,
limiter, or source-discretization effects.

\subsection{Geometry-Rest Preservation Test}
\label{subsec:rest-state-drift}

The continuous selected source-balanced law admits the geometry-rest state
$a_i=R_0(z_i)^2$, $q_i=0$ under compatible zero-flow boundary data
(Proposition~\ref{prop:selected-rest-state-compatibility}). The current
principal implementation is not well balanced for that state. The
Rusanov/MUSCL operator includes numerical viscosity in the mass flux, and for
spatially varying $R_0$ this produces a nonzero discrete residual unless an
equilibrium reconstruction or path-conservative flux is added.

The zero-forcing rest runner starts the selected C23 and C40 geometries from the
rest state with a zero-velocity inlet and reports
\begin{flalign*}
    \quad
    &\max_i |q_i|,&
    &\max_i |a_i-R_0(z_i)^2|,&
    &\left|\sum_i a_i\Delta z-\sum_i a_i(0)\Delta z\right| &&
\end{flalign*}
as functions of grid size and elapsed time. This test is required context for
any comparison run initialized from the same geometry-rest state.

\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
bin/stenosis-hemodynamics verify rest \
  --severities 23,40 \
  --nxs 100,200,400,800 \
  --elapsed-times 0,0.001,0.01,0.1,1.0 \
  --output-dir report/assets/tables/verification \
  --overwrite
\end{lstlisting}

\IfFileExists{report/assets/tables/verification/rest_state_drift.tex}{%
    \begin{center}
    \fbox{\begin{minipage}{0.92\linewidth}
    \small
    \textbf{Principal numerical limitation.} The zero-forcing geometry-rest
    runs retain artificial solver flow at $t=1$ s on the production
    comparison-flow scale. This table is therefore the chapter's main negative
    numerical result.
    \end{minipage}}
    \end{center}
    \input{report/assets/tables/verification/rest_state_drift.tex}
}{%
    \begin{center}
    \small
    \emph{Rest-state drift table unavailable in this checkout. Run
    \texttt{verify rest} to populate
    \path{report/assets/tables/verification/}.}
    \end{center}
}

The corrected rest-state run has zero body forcing and zero inlet flow, so the
reported motion is not an imposed boundary through-flow. The production
comparison-flow scale is
$q_{\mathrm{comp}}\approx2.288/\pi=0.7283\,\mathrm{cm}^3/\mathrm{s}$ in the
solver coordinate, corresponding to
$Q_{\mathrm{comp}}\approx2.288\,\mathrm{cm}^3/\mathrm{s}$ in physical circular
flow, as used in the axial-flow comparison
(Figure~\ref{fig:t1-axial-flow-comparison}). On the finest grid, the
$t=1$ s zero-forcing rest-flow values are
$0.7313$ and $0.7324\,\mathrm{cm}^3/\mathrm{s}$ in solver $q$ for C23 and C40,
with physical flows $\pi q=2.297$ and $2.301\,\mathrm{cm}^3/\mathrm{s}$. These
are the same order as the production comparison-flow scale, and they are
numerically close to it. Peak rest drift occurs earlier at $t=0.01$ s, reaching
$1.566$ and $1.658\,\mathrm{cm}^3/\mathrm{s}$ in solver $q$ on the finest C23
and C40 rows.

This geometry-rest failure is the principal numerical result of the chapter.
The selected C23/C40 comparison runs are initialized from the same geometry-rest
state and use the same operator, so their start-up transient may include
artificial rest-state motion before and during the imposed through-flow. The
available C23/C40 comparison therefore remains a velocity-output diagnostic
under the declared observation operator, not a clean discretization-accuracy
study.

\subsection{Boundary, CFL, Positivity, and Conservation Diagnostics}
\label{subsec:boundary-diagnostics}

The native solver records accepted timestep extrema, realized CFL extrema,
$\lambda_-$ and $\lambda_+$ extrema, the minimum subcriticality margin
$\min(-\lambda_-,\lambda_+)$, mass extrema and mass defect, positivity
projection count, and total area correction from the positivity floor. The
fixed-area characteristic boundary rule is used under the subcritical sign
condition $\lambda_-<0<\lambda_+$. A positive characteristic radicand alone is
not the boundary-regime evidence; the persisted sign extrema and subcriticality
margin are the relevant diagnostics.

In the rest-state table, the finest-grid C23/C40 $t=1$ s rows retain
positive subcritical margins and record the mass defects shown in the table, and the generated rest CSV records zero positivity projection counts for
these rest tests. The production C23/C40
comparison diagnostics in Table~\ref{tab:t1-alpha-radicand-diagnostics} provide
the corresponding characteristic-speed evidence for the runs later used in the
velocity-output comparison. These diagnostics are admissibility and bookkeeping
checks. They do not remove the non-well-balanced geometry-rest drift reported
above.

\subsection{Secondary Implementation-Health Checks}
\label{subsec:verification-summary}

The fourth evidence tier is kept out of the main argument. Appendix~\ref{app:num-secondary-implementation-health}
retains the package benchmark stage summary, sampled self-convergence figure,
backend-parity figure, nonselected DG p/h demonstration, resolved-velocity
benchmark stage, and broad rheology/profile sensitivity figure. Appendix~\ref{app:num-stationary-stokes-refinement}
retains the fixed-wall stationary-Stokes projection and mesh-refinement record.
Those records are useful implementation-health checks: they can reveal runnable
surface regressions, boundedness problems, backend-policy discrepancies, or
sensitivity to nonprincipal descriptors. They do not replace the MMS record, and
they do not change the geometry-rest conclusion.

\FloatBarrier
