\subsubsection{3D Fixed-Wall and Resolved-FSI Models}
\label{subsubsec:3d-models}

A fixed-wall 3D hemodynamic model solves for the velocity field
$\vu(t,\vx)$, pressure field $p(t,\vx)$, and stress-derived outputs on a
three-dimensional lumen with fixed or prescribed wall geometry
\cite{QuarteroniFormaggia2004CardiovascularModeling}. It can supply resolved
velocity, pressure, and wall traction. A WSS output additionally requires the
stress tensor, wall normal, wall location, tangent convention, and output
operator to be declared
\cite{JohnEtAl2017WallShearStressVectorForm}. These quantities are fields on
the resolved lumen and describe the vessel wall's interaction with the fluid
domain, not the 1D state variables $A$, $Q$, and $p_{\mathrm g}$.

Resolved FSI changes the wall descriptor. A fixed-wall model belongs to the
$\mathcal W_{\mathrm{rigid}}$ or prescribed-wall reading in
Definition~\ref{def:wall-fsi-closure-catalog}; a resolved-FSI model uses
$\mathcal W_{\mathrm{resolvedFSI}}$ and adds wall displacement, wall velocity,
structural stress, and fluid--structure interface conditions. The elastic 1D
wall law $\mathcal W_{\mathrm{elastic1D}}$ is a reduced wall closure, not a
structural FSI solve
\cite{FormaggiaEtAl2007FSICoupling}.

The 3D/FSI tier can represent eccentricity, curvature, branches, secondary
flow, separation, and recirculation more directly than a 1D area--flow model.
That role does not make it assumption-free. A valid comparison requires matched
geometry, rheology descriptor $\mathcal R$, wall closure $\mathcal W$, boundary
data, mesh and time resolution, output operators, and discrepancy metrics. A
velocity-slab diagnostic, a traction-based WSS comparison, and a pressure-ratio
comparison are different observation problems.

The 3D/FSI tier supplies the reference vocabulary for resolved velocity,
pressure, traction, and wall motion, but it does not automatically validate a
reduced model without matched comparison data. It is not the selected broad
parametric generator in this report because the setup and computational burden
are high, not because resolved 3D or FSI models are unimportant.
