\subsubsection{Multiscale Coupling}
\label{subsubsec:multiscale-models}

A multiscale model is a coupled forward problem that combines tiers of the
hierarchy. A concrete 3D--1D--0D example is a resolved local 3D stenosis segment
embedded between upstream and downstream 1D arterial vessels, with terminal
0D RCR or Windkessel beds closing the distal circulation. The purpose is to keep
local resolved physics where it matters while representing pulse-wave
propagation and distal impedance without making the entire circulation a 3D
domain
\cite{QuarteroniVeneziani2003GeometricalMultiscale}.

The interface laws are part of the model. At a 3D--1D interface, the record must
state which 3D cross-section is coupled, how 3D velocity and traction are reduced
to 1D flow or mean pressure, and whether compatibility is imposed through
pressure, normal traction, total flow, characteristic variables, or a stabilized
coupling variant. At a 1D--0D interface, terminal 1D flow enters the ODE load
and the 0D circuit returns a pressure, impedance, or characteristic boundary
relation. Resolved-FSI coupling additionally requires compatible wall
kinematics, structural stresses, and interface work. These choices affect
stability and interpretation, and they must be recorded with the numerical
coupling convention
\cite{FormaggiaEtAl2007FSICoupling}.

A multiscale model is not a new fidelity tier; it is a coupled forward problem
whose interface laws are part of the model definition. Coupling more components
does not automatically make a comparison more correct. The boundary data,
interface variables, time coupling, output maps, and discrepancy metrics decide
what claim the computation can support. In this report, multiscale coupling
provides comparison and boundary-condition vocabulary; it is not the baseline
single-vessel solve unless explicitly declared.
