\section{Numerical Methods, Verification, Validation, and Observation Operators}
\label{sec:numerical-methods}
\label{sec:numerical-methods-review}

Numerical evidence for stenotic-flow models is inseparable from the method that
produces the state, the reference data used to check it, and the operator used to
observe it. A resolved incompressible-flow or FSI computation, a distributed 1D
area--flow network, and a diagnostic postprocessing map can all report velocity
or flow, but they do so through different discrete spaces, stability
constraints, conservation properties, and measurement conventions. This chapter
therefore reviews the numerical-method and evidence standards needed to interpret
the StenosisHemodynamics audit.

\subsection{Finite-Element Incompressible Flow and FSI}
\label{subsec:numerical-methods-fem-fsi}

Resolved hemodynamic models commonly start from incompressible Navier--Stokes or
Stokes equations, optionally coupled to a deforming wall. A finite-element
realization writes those equations in weak form, chooses discrete velocity and
pressure spaces, and solves the resulting saddle-point problem. The mixed
velocity--pressure pair is not a cosmetic choice: the discrete spaces must
control the pressure null mode and avoid spurious pressure oscillations. Stable
Taylor--Hood-type pairs, divergence-conforming constructions, or equal-order
pairs augmented by pressure-stabilization, streamline-upwind/Petrov--Galerkin,
least-squares, local projection, or grad-div terms are different numerical
contracts, even when they target the same continuum equations
\cite{GaldiEtAl2008HemodynamicalFlows}.

When the wall moves, the numerical contract also includes the fluid--structure
interface. FSI formulations must transfer traction and velocity or displacement
data across the interface, update the fluid domain or mesh, and control the
added-mass and time-lag effects that appear in partitioned coupling. Monolithic
and partitioned schemes can both be useful, but they expose different solver
errors, stability limits, and coupling residuals
\cite{FormaggiaEtAl2007FSICoupling,QuarteroniVeneziani2003GeometricalMultiscale}.
For a reduced 1D comparison, a resolved FSI or CFD field is therefore not a raw
truth object; it is a numerically produced state that still needs its own mesh,
timestep, boundary, material, and output-operator record.

\subsection{Finite Differences, Finite Volumes, DG, and High-Resolution Schemes}
\label{subsec:numerical-methods-balance-law-families}

One-dimensional blood-flow solvers are usually written as balance-law solvers.
Finite-difference schemes approximate derivatives at grid points and can be
compact and high-order, but conservation and geometric-source balance must be
checked at the discrete level. Finite-volume schemes instead advance cell
averages using numerical fluxes through cell faces. That structure makes local
conservation visible and is well suited to wave propagation, boundary fluxes,
and discontinuity-capturing limiters
\cite{LeVeque2002FiniteVolumeMethodsForHyperbolicProblems}. In arterial
networks, node conditions and terminal models add another layer: each vessel
solve must be compatible with coupling rules that determine incoming and
outgoing characteristic information \cite{Wang2014BloodFlowNetworks}.

Discontinuous Galerkin methods combine finite-volume fluxes with local
polynomial approximation. They can raise spatial order, expose element-local
residuals, and support $p$-refinement studies, but they still require numerical
fluxes, limiters or troubled-cell controls, and timestep choices. WENO and other
high-resolution reconstructions serve a related purpose for finite-volume and
finite-difference methods: they recover high-order accuracy on smooth regions
while suppressing nonphysical oscillations near steep gradients. In stenotic
area--flow problems, these design choices interact with wall and geometry
sources, so smooth-solution accuracy alone is not enough. A scheme can show good
manufactured-solution behavior and still fail to preserve a physiologically
meaningful rest state if flux and source terms are not balanced.

\subsection{Stability, Positivity, Conservation, and Well-Balancing}
\label{subsec:numerical-methods-stability-positivity-balance}

Explicit wave-propagation schemes carry CFL restrictions: the timestep must stay
compatible with the grid scale and characteristic speeds. Semi-implicit and
implicit network methods can relax the fastest explicit restriction, especially
when wave speed or network coupling is stiff, but they replace it with nonlinear
or linear solver tolerances, splitting errors, and coupling-residual criteria.
Open and benchmarked 1D hemodynamics workflows therefore record not only the
equation family, but also the timestep policy, boundary assumptions, network
coupling, and diagnostic metadata
\cite{BoileauEtAl2015Benchmark,BenemeritoEtAl2024OpenBF,BeckersKolbe2025LaxFriedrichsHemodynamics,
LuccaEtAl2025FFRPrediction}.

For compliant-vessel balance laws, three checks are especially direct. First,
positivity guards ensure that area or radius states remain physical; any
projection or floor must be counted because it changes the discrete solution.
Second, conservation diagnostics record whether mass or flow bookkeeping is
consistent with boundary fluxes and source terms. Third, well-balanced tests
ask whether a target equilibrium is preserved by the discrete operator itself.
High-order well-balanced schemes make this requirement explicit by reconstructing
or integrating flux and source contributions so that selected rest states cancel
after discretization, not only in the continuous equations
\cite{PimentelGarciaEtAl2025HighOrderWellBalanced}. For the present report,
geometry-rest preservation is therefore a core numerical property rather than a
postprocessing preference.

\subsection{Verification, Validation, and Reference Data}
\label{subsec:numerical-methods-verification-validation}

Verification asks whether the implementation solves the equations it claims to
solve. Manufactured solutions are useful because the analyst chooses a smooth
state, derives the forcing needed to make that state exact, and then checks the
discrete error under grid and timestep refinement. Exact or manufactured states
can probe fluxes, sources, time integration, boundary traces, and coordinate
maps without requiring physiological data. They also expose whether a nominally
high-order method is limited by boundaries, source quadrature, limiters, or
operator splitting.

Grid and timestep convergence studies then ask whether changes in resolution
change the reported quantities in the expected direction and magnitude. These
studies are strongest when spatial and temporal effects are separated; otherwise
they are better reported as sensitivity or time-step-insensitivity checks.
Benchmark problems provide a broader reproducibility context by comparing
multiple solvers on shared model cases, but a benchmark row does not by itself
prove that a different stenosis geometry, wall law, or boundary rule is
accurate \cite{BoileauEtAl2015Benchmark}.

Open modeling platforms and model repositories occupy a related but distinct
evidence role. A platform such as SimVascular supports inspectable model
construction, meshing, simulation, and postprocessing workflows, while a
vascular model repository provides reusable geometry and model records
\cite{SimVascular2026Platform,SimVascular2026VascularModelRepository}. Those
resources strengthen reproducibility and make reference cases easier to audit,
but they are not themselves validation evidence unless the geometry, boundary
data, material assumptions, numerical settings, and observable are matched to
the claim being tested.

Validation is a different evidence category. It requires comparison with an
accepted reference for the target question, such as controlled experimental
measurements, clinical measurements with known acquisition conventions, or a
resolved computational reference whose geometry, material parameters, boundary
history, timestep, and observable match the reduced model closely enough to
support the intended claim. Computational reference data are often the practical
first tier for method development, while experimental data are needed when the
claim concerns physical or physiological prediction. In both cases, the
observation operator determines what quantity is actually being compared.

\subsection{Cross-Dimensional Observation Operators}
\label{subsec:numerical-methods-observation-operators}

Reduced and resolved models do not naturally live in the same data space. A 1D
area--flow state supplies section area, flow, and mean axial velocity after a
coordinate convention is fixed. A 3D velocity field supplies nodal, elemental, or
quadrature-point velocities on a mesh. Comparing them therefore requires an
observation operator: a declared map from each numerical state to a common
observable.

For section comparisons, the operator must specify the target planes, mesh
intersection rule, interpolation of velocity onto the cut geometry, quadrature
weights, area normalization, and treatment of empty or invalid cuts. For profile
comparisons, it must also specify the radial coordinate, binning rule, profile
closure used by the 1D reconstruction, and any azimuthal averaging. Multiscale
and FSI literature treats such maps as part of the model interface rather than
as after-the-fact plotting choices
\cite{FormaggiaEtAl2007FSICoupling,QuarteroniVeneziani2003GeometricalMultiscale}.
Clinical FFR and CT-FFR sources show the same requirement in diagnostic form:
the reported pressure-ratio observable is meaningful only after the imaging,
model, boundary, and pressure-extraction conventions have been fixed
\cite{TebaldiEtAl2015FFROverview,NorgaardEtAl2014NXTFFRCT,ShaikhEtAl2025CTFFR}.
Comparisons of 1D and 3D FFR estimates make the cross-dimensional version of
that requirement explicit \cite{Blanco2018FFR1D3D}.
In this report, the plane--tetrahedron quadrature operator plays that role for
the available resolved 3D velocity data, while the 1D state is observed through
physical flow and area-mean velocity.

\subsection{Method Evidence Standards for the Present Audit}
\label{subsec:numerical-methods-evidence-standards}

The review above gives a compact standard for reading the numerical record.
Each reported result should identify its equation tier, discrete operator,
boundary rule, timestep policy, stability and positivity diagnostics, reference
data, and observation operator. Table~\ref{tab:numerical-methods-evidence-map}
summarizes the roles used by the present manuscript.

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Numerical-method evidence roles used to interpret the implemented
    stenosis audit.}
    \label{tab:numerical-methods-evidence-map}
    \begin{tabularx}{\textwidth}{@{}p{0.23\textwidth}p{0.31\textwidth}X@{}}
        \toprule
        \textbf{Evidence item} & \textbf{Numerical question} & \textbf{Role in this report} \\
        \midrule
        Manufactured solution
            & Does the implemented forced operator reproduce a declared smooth exact
            state under refinement?
            & Implementation-verification evidence for flux, source, boundary, and
            time-integration wiring; strongest when the manufactured residual is
            audited outside the production right-hand-side path or by mutation
            tests. \\
        Geometry-rest equilibrium
            & Does the discrete flux/source operator preserve a selected rest state?
            & Well-balanced-property test for the stenosis geometry and wall-source
            realization. \\
        CFL, positivity, conservation
            & Did the accepted run remain inside the recorded admissible numerical
            regime?
            & Diagnostic bookkeeping for timestep, characteristic, area, and mass
            behavior. \\
        Benchmarks and reference data
            & How does a solver compare on shared cases or against accepted data?
            & Reproducibility and validation context, conditional on matched model
            and observable assumptions. \\
        Plane--tetrahedron quadrature
            & What common observable is extracted from resolved 3D and 1D states?
            & Cross-dimensional observation operator for section velocity, flow, and
            radial-profile diagnostics. \\
        \bottomrule
    \end{tabularx}
\end{table}

\begin{stenosisillustration}
For the StenosisHemodynamics evidence record, the three central checks map
directly onto the review concepts above. The MMS run is the manufactured-state
verification problem: it checks whether the declared finite-volume operator,
forcing residuals, boundary traces, and timestepper reproduce a known smooth
area--flow solution. The rest-equilibrium run is the well-balanced test: it asks
whether the discrete stenosis geometry, wall source, and numerical flux preserve
the quiescent state that the continuous balance law admits. The
plane--tetrahedron quadrature rule is the observation-operator test: it defines
how a resolved 3D velocity field becomes section area, physical flow, mean axial
velocity, and radial-bin data comparable to 1D outputs. These checks answer
different numerical questions, so agreement in one cannot substitute for
a limitation or ambiguity in another.
\end{stenosisillustration}
