\section{Integrated Discussion}
\label{sec:discussion}

\subsection{RQ1: Model Hierarchy and Question Matching}
\label{subsec:discussion-model-hierarchy-question-matching}

The first research question asked how blood-flow model classes are organized
and what distinguishes them. The review supports the conclusion that the
hierarchy is a set of retained-state, closure, coupling, and observation
choices, not a ranking of accuracy. Formaggia, Lamponi, and Quarteroni show
that section-averaging and wall simplification define different 1D operators;
Shi, Lawford, and Hose distinguish lumped global pressure--flow interactions
from distributed 1D pulse propagation; and Quarteroni--Veneziani multiscale
coupling shows that boundary data can be a reduced model of omitted circulation
\cite{FormaggiaEtAl2003OneDimensionalBloodFlow,ShiEtAl2011BloodFlowModelReview,QuarteroniVeneziani2003GeometricalMultiscale}.
Thus 0D, 1D, and 2D formulations should not be described as weaker versions of
3D CFD; they retain different states and require different observation operators.

The case study demonstrates this answer by using a distributed 1D
area--flow state for quantities it can natively or reconstructively support:
area, flow, section-mean velocity, a wall-law pressure representative, and a
specified velocity observation operator. Native WSS, secondary flow, wall
stress, and clinical pressure-ratio evidence require different states or
observations.

\subsection{RQ2: Observables and Closure Choices}
\label{subsec:discussion-closures-interpretability}

The second research question asked how rheology, wall mechanics, geometry,
boundary data, numerical discretization, evidence standards, and observation
operators affect interpretation. These choices define the quantity being
reported. Blanco and coauthors support this point for
FFR-specific 1D--3D comparison under matched pressure boundary conditions and
dedicated loss closures; Liu and coauthors show observable-dependent rheology
sensitivity in an intracranial stenosis setting; John and coauthors show that
WSS-vector convention changes AWSS/OSI; and Gambaruto and coauthors show that
geometry reconstruction can be a material sensitivity source in a patient
aneurysm case
\cite{Blanco2018FFR1D3D,LiuEtAl2021ICASRheology,JohnEtAl2017WallShearStressVectorForm,GambarutoEtAl2011GeometryRheologySensitivity}.

The case study therefore presents its closure list as mathematical information:
Newtonian principal rheology, parabolic profile closure, an elastic
pressure--area wall law, analytic reference geometry, fixed-area characteristic
boundary approximation, and a plane--tetrahedron velocity observation operator.
The resulting comparison is a velocity-observation comparison under those
assumptions; WSS, wall-displacement, and FFR claims require separate observable
definitions and evidence.
The membrane discussion sharpens the same boundary: Newtonian or
generalized-Newtonian terminology belongs to the fluid stress law, while the
membrane model is a structural wall closure and the quasi-static comparator
provides wall-motion scale rather than full transient moving-boundary FSI
evidence.

\subsection{RQ3: Numerical Evidence and Cross-Dimensional Interpretation}
\label{subsec:discussion-numerical-evidence-observation}

The third research question asked what limitations remain when resolved scale,
cost, identifiability, validation status, reproducibility, and uncertainty are
compared. The review supports the conclusion that interpretation depends on the
specified equation, state variables, discretization, output map, and metric, and
each evidence type checks only part of that chain.
ASME VVUQ terminology separates verification, validation, and uncertainty
quantification; Boileau and coauthors provide benchmark comparison evidence;
Pimentel-Garcia and coauthors show that a well-balanced claim must name the
preserved stationary family; and Colebank shows that practical identifiability
depends on measured outputs, boundary-condition class, and design assumptions
\cite{ASME2022VVUQTerminology,BoileauEtAl2015Benchmark,PimentelGarciaEtAl2023HighOrderWellBalanced,Colebank2025HemodynamicsIdentifiability}.

The worked stenosis example demonstrates the distinction. Its manufactured
solution checks one implemented forced operator using shared manufactured
states and lower-level operator utilities. Its geometry-rest test addresses
equilibrium-preservation testing for the chosen discretization and boundary
approximation; the selected geometry-rest-balanced finite-volume realization
preserves the named rest family to roundoff, while the historical MUSCL/Rusanov
row remains useful as nonselected context. Its 1D--3D comparison becomes
meaningful only after both models are reduced to the same section-mean velocity
observable and discrepancy metric. These checks support a bounded diagnostic
cross-model comparison. The positive MMS result, rest-preservation result, and
final-time comparator are useful precisely because they belong to different
parts of the same interpretive chain rather than to one validation grade.

\subsection{Contribution and Limitations}
\label{subsec:discussion-contribution-limitations}

The contribution of this report is methodological rather than clinical. It
offers a way to read mathematical hemodynamics across model tiers by keeping
state variables, closures, boundary data, discretization, observation rules,
metrics, and evidence types in view at the same time. The idealized stenosis
case study matters because it does not merely restate that principle
abstractly; it carries it through one specified model and shows how different
kinds of numerical evidence sit alongside one another. In that sense, the
positive MMS result, the geometry-rest preservation result, and the bounded
1D--3D comparison are jointly informative.

The revised comparison also changes the numerical interpretation. The selected
geometry-rest-balanced method removes a numerical confound that previously made
production-grid rest drift comparable in scale to the observed 1D--3D
discrepancy. The remaining discrepancy is therefore better read as a
conditional output of the model, closure, boundary, and observation chain
rather than as a mixture dominated by a known geometry-rest defect. Because the
resolved comparator still lacks matched transient boundary and wall histories,
the result remains a diagnostic comparison rather than model-form validation.

The limitations follow from the same discipline. The review is narrative and
scoping rather than systematic. Some sources remain preprints, software
artifacts, or metadata entries with roles narrower than peer-reviewed empirical
evidence. The case geometry, wall law, rheology, and boundary approximation are
idealized. The comparison data are informative but incomplete, especially for
transient wall and boundary matching. The reproducibility materials are strong
while stopping short of a fully archived public release with packaged raw 3D
inputs. These limitations do not diminish the report's main contribution; they
define it accurately.

\section{Conclusion}
\label{sec:conclusions-limitations}

The main conclusion of this report is that blood-flow simulation results cannot
be read responsibly from model dimension alone. What matters is the full
interpretive setting: which state is retained, which physics are closed, what
data are supplied, how numerical outputs are observed, and what kind of
evidence is available for the claim being made. Read in that way, 3D CFD, FSI,
lower-dimensional models, multiscale systems, and learned surrogates are not
simply points on a single accuracy ladder. They are different mathematical
descriptions whose outputs become comparable only under specified observables
and evidence standards.

The idealized stenosis case study shows what that discipline looks like in
practice. For the reported 1D realization, the manuscript supports
implementation-verification evidence for the forced operator, records
geometry-rest preservation for the selected finite-volume comparison method,
and reports final-time section-mean velocity discrepancies under a specified
1D--3D observation map. Those findings are meaningful precisely because they
are bounded. The older non-well-balanced MUSCL/Rusanov row remains a useful
nonselected diagnostic, but it is no longer the numerical method behind the
reported comparison rows. Future work should strengthen the same chain where
the present evidence remains limited: obtain matched wall and boundary metadata
for resolved comparisons, report uncertainty separately from model discrepancy,
and specify validation observables before predictive or clinical claims are
made.
