\section{Literature Synthesis and Open Challenges}
\label{sec:literature-synthesis}
\label{sec:literature-synthesis-open-challenges}

The preceding chapters introduced the review axes one at a time: continuum
fields, model hierarchy, closure choices, observables, numerical methods, and
evidence standards. This chapter compares across those axes. Its purpose is not
to add another catalog of sources, but to state what the landscape collectively
shows. The central finding is that blood-flow models are interpretable only
when retained state, resolved scale, data, closure choices, observables,
numerical evidence, and validation status are aligned with the question being
asked.

The synthesis can be organized into four proposition families used throughout
the remainder of the report:
\begin{enumerate}
    \item \textbf{Model state.} Model hierarchy is a hierarchy of retained
    state and observation maps, not a monotone hierarchy of accuracy; dimension
    reduction changes closure requirements and native observables, not merely
    computational cost.
    \item \textbf{Closures and observables.} Local stenosis hemodynamics depends
    on omitted circulation through boundary or multiscale coupling models, and
    rheology, wall, and geometry sensitivity must be reported relative to a
    specified observable and regime.
    \item \textbf{Evidence and validation.} Agreement between 1D and 3D models
    is interpretable only under matched model specifications and a common observation
    operator; manufactured solutions, grid studies, equilibrium tests,
    benchmarks, cross-model comparisons, external validation, UQ, and
    reproducibility support different claims.
    \item \textbf{Learned and identifiable maps.} Practical identifiability
    depends on measured outputs, boundary model, noise assumptions, and
    experimental design; learned models inherit the geometry family, boundary
    inputs, source-model assumptions, training distribution, and target
    observable of their data.
\end{enumerate}

\subsection{Resolved Scale, Cost, and Retained State}
\label{subsec:synthesis-fidelity-identifiability-observables}

Increasing resolved scale can expose local velocity gradients, wall
interaction, and geometry-specific flow features, but it also increases the
number of boundary, material, geometric, and numerical assumptions that must be
identified. The continuum and FSI sources show what is gained by retaining
local flow fields, wall kinematics, and interface work; the multiscale sources
show that those fields still have to be connected to lower-dimensional
circulation states; Formaggia, Lamponi, and Quarteroni show that structural
simplifications change the one-dimensional differential operator itself; and
the 0D/1D review literature shows what is deliberately compressed into axial
or lumped pressure--flow variables
\cite{FormaggiaEtAl2003OneDimensionalBloodFlow,FormaggiaEtAl2007FSICoupling,QuarteroniVeneziani2003GeometricalMultiscale,ShiEtAl2011BloodFlowModelReview}.
The reduced tiers are not weaker by definition. They are appropriate when the
question concerns pulse propagation, network pressure--flow behavior, rapid
screening, or parameter studies for which fully resolved data and boundary
conditions are unavailable or unnecessary.

The cost of that compression is interpretive. A 1D model may have fewer state
variables than a 3D model, but it can still be underdetermined if wall, profile,
friction, rheology, and boundary closures are treated as adjustable without
independent evidence. Conversely, a highly resolved forward model can depend on
material properties, prestress, inlet histories, outflow impedances, geometry
state, and wall motion that are rarely all observed with matching certainty.
Parameter-identifiability work for hemodynamics PDE models makes this tension
explicit: changing measured outputs and boundary-condition class changed which
parameters could be inferred in a 1D pulmonary model
\cite{Colebank2025HemodynamicsIdentifiability}. Adding modeling detail is
therefore valuable only when the data and outputs can distinguish the resulting
parameters.
The synthesis point is therefore comparative rather than taxonomic: the right
tier is the one whose retained state is identifiable for the claim being made.

\subsection{Observables and Cross-Model Comparability}
\label{subsec:synthesis-observables-comparability}

Observation operators are part of the model, not post-processing decoration. A
1D area--flow solution does not natively contain the same observable as a
resolved 3D velocity field, and a pressure-drop, flow-rate, section-mean
velocity, wall-shear, or FFR-style quantity each imposes a different comparison
setup. Multiscale coupling work already treats interface quantities and
compatibility conditions as mathematical structure
\cite{FormaggiaEtAl2007FSICoupling,QuarteroniVeneziani2003GeometricalMultiscale};
diagnostic 1D--3D comparison needs the same discipline at the output side. A
flow-rate comparison, for example, tests a different map than a pressure-ratio
or wall-shear comparison even when the underlying geometry is shared.

The clinical FFR literature sharpens the point because it separates anatomical
stenosis appearance from pressure-flow function: Tonino and coauthors supply
the angiographic-versus-functional contrast, while FFR overview work defines
the scalar as a pressure-ratio diagnostic
\cite{ToninoEtAl2010FAMEAngioFunctional,TebaldiEtAl2015FFROverview}.
CT-derived FFR and 1D--3D FFR comparison sources then move that diagnostic into
computational chains. Blanco and coauthors compared 20 patients and 29
geometries under matched pressure boundary conditions and dedicated 1D stenosis
and bifurcation loss terms, so their result supports an FFR-specific
matched-observable comparison rather than field-level or tier-level equivalence
\cite{Blanco2018FFR1D3D}. Together with CT-FFR methodology work, this shows
that the scalar is not a raw geometry measurement; it is an output of an
acquisition, modeling, boundary, and pressure-observation chain
\cite{NorgaardEtAl2014NXTFFRCT,ShaikhEtAl2025CTFFR}. Thus a cross-model
discrepancy is uninterpretable until the observed quantity is declared and
matched.

\subsection{Uncertainty in Boundaries, Walls, Rheology, and Geometry}
\label{subsec:synthesis-closure-geometry-uncertainty}

Boundary-condition uncertainty is one of the most persistent barriers to
interpretable hemodynamics. Inlet waveforms, outlet impedances, terminal
resistances, and coupled-domain histories can change the local solution as much
as the internal discretization. In a multiscale setting, these quantities are
not auxiliary inputs; they are the mechanism by which upstream and downstream
vascular beds communicate with the modeled region
\cite{QuarteroniVeneziani2003GeometricalMultiscale,FormaggiaEtAl2007FSICoupling}.
A reduced stenosis result that does not declare its boundary approximation
therefore cannot be cleanly compared with a resolved result, even if both
models share a nominal geometry. The source contrast is direct: multiscale
papers treat the boundary as an interface model, whereas single-domain
diagnostic studies can hide the same choice behind an inlet or outlet label.

Wall and rheology uncertainty create a second closure layer. Wall stiffness and
thickness determine wave speed and area response; rheology and velocity-profile
assumptions determine dissipative behavior and momentum transport. In 1D
models, these effects enter through pressure--area laws, friction terms,
effective viscosity choices, and momentum-flux correction factors
\cite{ShiEtAl2011BloodFlowModelReview}. In FSI models, they enter through wall
material laws, structural boundary conditions, and the coupling algorithm. Both
settings face the same interpretive problem: when the wall or rheology is not
independently constrained, a discrepancy can be a physics signal, a closure
signal, or a compensation between uncertain choices.

These uncertainties are not of one kind. Aleatory variability concerns
physiological variation across cycles or populations; epistemic or model-form
uncertainty concerns missing mechanisms; parameter uncertainty concerns unknown
values in a selected model; numerical error concerns the computed realization;
and data-matching uncertainty concerns whether two models, measurements, or
files describe the same state. Liu and coauthors show why the category must be
tied to an observable: rheology choice was negligible for a pressure ratio in
their intracranial stenosis setting but visible in low-WSS regions
\cite{LiuEtAl2021ICASRheology}.

Geometry is similarly double-edged. Patient-specific geometry is attractive
because it carries clinically meaningful morphology, but segmentation,
registration, current-versus-reference configuration, and wall-motion state can
dominate the modeling specification. Idealized geometry removes many of those
ambiguities and exposes controlled severity and refinement parameters, but it
cannot by itself answer patient-specific questions. The useful distinction is
not realistic versus artificial. It is whether the geometry serves the claim:
patient-specific geometry for patient-specific interpretation, and idealized
geometry for isolating numerical, closure, and comparison behavior. Gambaruto
and coauthors provide a bounded example: in one aneurysm case, reconstruction
variation within imaging resolution changed WSS enough that geometry
sensitivity was greater than, but comparable to, rheology sensitivity
\cite{GambarutoEtAl2011GeometryRheologySensitivity}. That is not a universal
uncertainty range; it is evidence that reconstruction can be a claim-level
model input.

\subsection{Validation Status, Reproducibility, and Evidence Limits}
\label{subsec:synthesis-consistency-verification-validation}

Verification, validation, uncertainty quantification, and reproducibility
should not be collapsed. ASME VVUQ terminology is useful because it separates
conformity to a mathematical description, representation of a real application,
and propagation of uncertain numerical or physical quantities
\cite{ASME2022VVUQTerminology}. Manufactured solutions and grid studies ask
whether an implementation behaves as expected for a declared mathematical
problem. Benchmarks and open solvers
support reproducibility and comparison across workflows
\cite{BoileauEtAl2015Benchmark,BenemeritoEtAl2024OpenBF}. The benchmark study
contributes shared cases and scheme comparison; OpenBF contributes an
inspectable open finite-volume workflow. Resolved-flow platforms and model
repositories extend that reproducibility lesson to 3D workflows by making
geometry, meshing, solver, and repository materials easier to inspect
\cite{SimVascular2026Platform,SimVascular2026VascularModelRepository}. None of
these materials is automatically a validation claim unless the model, data,
observable, and accepted reference match the target question.

Equilibrium preservation is a particularly sharp consistency test. A continuous
stenotic balance law may admit a rest or steady state through cancellation
between flux gradients, geometry terms, wall response, friction, and boundary
data. A discrete method that does not preserve that state can generate motion
from the numerical representation itself. Well-balanced methods for 1D blood
flow make this point directly and raise the standard for stenosis
computations: manufactured solutions and grid studies are useful, but an
equilibrium problem needs an equilibrium-preservation check
\cite{PimentelGarciaEtAl2023HighOrderWellBalanced}. The 2023 Pimentel-Garcia
et al. claim concerns all stationary solutions for discontinuous geometrical
and mechanical properties, so its evidence role is equilibrium preservation
rather than generic high-order behavior. The broader verification and validation
gap is that many studies report plausible hemodynamic outputs without enough
evidence to separate implementation error, discretization error, closure
uncertainty, cross-model mismatch, and external validation against an accepted
target.

\subsection{Reproducibility and Emerging Learned Methods}
\label{subsec:synthesis-reproducibility-learned-methods}

Reproducibility is the practical bridge between modeling ambition and reviewable
evidence. Benchmark studies help by fixing common cases and comparison
quantities, but stenosis studies also need persisted geometry definitions,
boundary histories, closure parameters, solver settings, diagnostics, and output
maps \cite{BoileauEtAl2015Benchmark,BenemeritoEtAl2024OpenBF}. The lesson from
the benchmark/open-solver pair is not that every stenosis study must use the
same code. It is that disagreement between two models is hard to assign to
physics, parameters, numerics, or data handling unless the comparison design is
inspectable.

Learned methods do not remove this requirement. Physics-informed neural networks
offer a framework for solving and inverting PDE-constrained problems
\cite{RaissiEtAl2019PhysicsInformedNeuralNetworks}, and hemodynamics reviews
now treat PINNs as an active direction for vascular modeling
\cite{YuEtAl2026PINNHemodynamicsReview}. At the same time, gradient-pathology
analyses show that training dynamics can be a limiting numerical issue rather
than a transparent solver detail
\cite{WangEtAl2021PINNGradientPathologies}. Operator-learning methods such as
DeepONet and Fourier neural operators shift the ambition from one solution to a
map between function spaces \cite{LuEtAl2021DeepONet,LiEtAl2021FNO}. Recent
hemodynamics-specific neural differential and stenosed-vessel operator-learning
work shows why such models are appealing for fast parameterized flow prediction.
Csala and coauthors report lower error than a specified 1D FEM baseline and
sub-1.2\% relative error against averaged 3D data within their declared
train/test setting \cite{CsalaEtAl2025PCNDE}. Velikorodny and coauthors learn
pressure and flow maps and then derive FFR in a simplified stenosis setting
with 201 training and 68 test cases
\cite{VelikorodnyEtAl2025DeepOperatorStenosed}. These are conditional
dataset-level claims, not evidence of patient-specific or out-of-distribution
clinical validity. Their open challenge is the same synthesis problem in a new
form: the training distribution, geometry family, boundary inputs, wall and
rheology assumptions, loss normalization, and validation observables must be
declared before speed or fit can be interpreted as hemodynamic evidence.

\subsection{Role of the Idealized Stenosis Case Study}
\label{subsec:synthesis-idealized-stenosis-case-study}

The idealized stenosis case study is useful because it carries one declared
equation set, retained state, closure list, boundary approximation,
discretization, observation rule, and metric definition through one controlled
example. Within that setting it places three bounded statements side by side:
positive manufactured-solution verification, a negative geometry-rest
preservation result, and a constrained 1D--3D velocity comparison. It removes
segmentation and clinical-matching ambiguity so that closure dependence,
boundary sensitivity, equilibrium preservation, verification evidence, and
cross-dimensional observability can be examined under one declared comparison
setup. Patient-specific interpretation remains the domain of studies with
matched patient geometry, boundary data, wall state, and external validation
data.
