\section{Background and Literature Review}
\label{sec:background-literature-review}

This chapter reviews the modeling and verification context needed for the
reported study. It deliberately avoids re-deriving continuum mechanics in the
main text. The required continuum assumptions, transport identities, and
governing-equation details are available in the appendices; the review here
instead compares model tiers, closure choices, verification practice, and
one-dimensional to three-dimensional observation operators.

\subsection{Stenotic Hemodynamics and Relevant Observables}
\label{subsec:stenotic-hemodynamics-observables}

Stenosis is a geometric and hemodynamic problem: a local lumen reduction can
alter axial velocity, pressure drop, wall loading, recirculation, and the
relationship between flow and downstream perfusion. Clinical and computational
literatures therefore use different observables depending on the question.
Pressure-ratio work emphasizes pressure drops and fractional-flow style
indices \cite{PijlsEtAl1996FFRFunctionalSeverity,ToninoEtAl2010FAMEAngioFunctional,TebaldiEtAl2015FFROverview},
whereas computational fluid-dynamics studies can expose local velocity,
traction, and wall-shear structures that are not section-averaged quantities
\cite{JohnEtAl2017WallShearStressVectorForm,GaldiEtAl2008HemodynamicalFlows}.
The present report uses velocity observables only. It does not report pressure
drop, FFR, CT-FFR, wall shear stress, or clinical validation.

Reduced models are attractive because stenosis studies often require repeated
parameter sweeps, reproducible solver comparisons, or network-level coupling.
The cost is that the relevant output must be compatible with the reduced state.
For a one-dimensional area--flow model, section mean velocity and physical flow
are natural; resolved radial profile, secondary flow, skewness, and wall
traction are not native outputs unless an additional reconstruction or
observation map is declared.

\subsection{Hierarchy of Computational Models}
\label{subsec:model-hierarchy-literature}

Three-dimensional CFD and FSI models retain geometry-resolved velocity and
stress fields and can represent local stenotic acceleration, separation, and
wall interaction more directly than reduced models
\cite{CourchaineRugonyi2018CardiacDevelopmentCFD,DaveyEtAl2024CardiacFluidDynamicsHeart}.
They also require detailed geometry, material assumptions, boundary histories,
mesh generation, and substantially higher computational cost. Coupled 3D--1D
or geometrical multiscale formulations address domain-size and boundary
limitations by joining local resolved regions with reduced vascular networks
\cite{FormaggiaEtAl2007FSICoupling,QuarteroniVeneziani2003GeometricalMultiscale,KopplHelmig2023DimensionReducedModeling}.

One-dimensional distributed models retain axial propagation and
pressure--area--flow dynamics while replacing the transverse velocity field by
closure terms \cite{FormaggiaEtAl2003OneDimensionalBloodFlow,QuarteroniFormaggia2004CardiovascularModeling,ShiEtAl2011BloodFlowModelReview}.
Zero-dimensional networks and terminal models are still cheaper and are useful
for lumped impedance and perfusion effects, but they do not resolve axial
stenosis localization. Intermediate 2D or axisymmetric models can retain more
local structure than 1D systems with less cost than full 3D, but they still
require their own symmetry and boundary assumptions.

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Model-tier roles relevant to the present report. The listed
    observables are admissible outputs only after the corresponding model and
    observation map have been declared.}
    \label{tab:model-tier-review}
    \begin{tabularx}{\textwidth}{@{}p{0.19\textwidth}p{0.22\textwidth}p{0.22\textwidth}X@{}}
        \toprule
        \textbf{Tier} & \textbf{Retained physics} & \textbf{Data and cost} & \textbf{Typical admissible outputs} \\
        \midrule
        3D CFD / FSI
            & Resolved velocity, local gradients, geometry, and possibly wall motion.
            & Detailed mesh, material and boundary histories; high cost.
            & Local velocity, pressure, traction, wall shear stress, section averages. \\
        Axisymmetric / 2D
            & More local structure than 1D under symmetry or planar assumptions.
            & Intermediate geometry and mesh burden.
            & Axial/radial fields consistent with the symmetry assumption. \\
        1D distributed
            & Axial wave propagation, area--flow dynamics, wall and friction closures.
            & Vessel centerline, radius, wall law, terminal and inlet data; low cost.
            & Area, flow, section mean velocity, reduced pressure variables. \\
        0D network
            & Lumped resistance, compliance, inertance, and terminal coupling.
            & Lowest spatial data burden; low cost.
            & Compartment pressures and flows, not local stenotic fields. \\
        Geometrical multiscale
            & Local resolved region coupled to reduced upstream/downstream networks.
            & Requires interface conditions and compatible observables.
            & Mixed local and network outputs with interface consistency checks. \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{One-Dimensional Models for Nonuniform and Stenotic Vessels}
\label{subsec:literature-1d-stenotic-vessels}

The defining step in a 1D vessel model is cross-sectional averaging. The state
typically stores a section area and volumetric flow, while the discarded
transverse structure re-enters through closure choices: momentum-flux factor,
velocity profile, wall law, friction, rheology, and source terms for nonuniform
reference geometry \cite{FormaggiaEtAl2003OneDimensionalBloodFlow,ShiEtAl2011BloodFlowModelReview}.
These choices are not interchangeable. A profile factor changes the convective
flux; a wall law changes the pressure--area relation and wave speed; a
friction or rheology law changes dissipative closure; and a stenosis-loss or
variable-radius term changes how geometry enters the reduced momentum balance.

Classical stenosis literature also shows that reduced pressure-loss formulas
and unsteady stenosis corrections can be useful but model a different closure
object than a distributed 1D finite-volume system
\cite{YoungTsai1973SteadyStenosis,YoungTsai1973UnsteadyStenosis,LyrasLee2021PressureDropROM}.
Recent stenosis-focused reduced models, including the Canic et al. extended 1D
formulation, add variable-radius contributions intended to better represent
stenotic geometry in a distributed balance
\cite{CanicEtAl2024Extended1DStenosis}. Coronary and arterial-tree models
make the same point at network scale: useful predictions depend on the
declared wall, terminal, and closure assumptions
\cite{DuanmuEtAl2019CoronaryArterialTree,Wang2014BloodFlowNetworks}.

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Closure dimensions that should be separated when comparing
    stenosis-aware 1D models.}
    \label{tab:one-dimensional-closure-review}
    \begin{tabularx}{\textwidth}{@{}p{0.20\textwidth}p{0.30\textwidth}X@{}}
        \toprule
        \textbf{Closure dimension} & \textbf{Common choices} & \textbf{Effect on evidence} \\
        \midrule
        State variables
            & Area--flow, radius--flow, pressure--flow, or network compartment variables.
            & Determines which outputs are native and which require reconstruction. \\
        Wall model
            & Fixed geometry, algebraic pressure--area relation, FSI coupling, terminal compliance.
            & Controls wave speed and pressure interpretation. \\
        Profile / momentum flux
            & Flat, parabolic, power-law, empirical correction factors.
            & Controls convective flux and any reconstructed radial profile. \\
        Friction / rheology
            & Newtonian, Carreau, Carreau--Yasuda, Casson, power-law, tube-flow laws.
            & Controls dissipation and shear-rate proxy interpretation
            \cite{PriesEtAl1992BloodViscosityTubeFlow,KumarEtAl2019CarreauYasudaHemodynamics,LiuEtAl2021ICASRheology}. \\
        Stenosis treatment
            & Reference-radius variation, empirical loss terms, variable-radius distributed terms.
            & Determines whether geometry appears only in area or also as a momentum source. \\
        Numerical method
            & Finite volume, discontinuous Galerkin, Lax--Friedrichs-type, method-specific boundaries.
            & Determines conservation, stability, and boundary evidence needs
            \cite{LeVeque2002FiniteVolumeMethodsForHyperbolicProblems,BeckersKolbe2025LaxFriedrichsHemodynamics}. \\
        Validation target
            & Benchmarks, manufactured solutions, 3D fields, pressure ratios, clinical measurements.
            & Determines whether the result is verification, cross-model comparison, or validation. \\
        \bottomrule
    \end{tabularx}
\end{table}

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Representative reduced-model and comparison references used to
    position the present implementation. The table is comparative rather than
    exhaustive; it records why each class is relevant and what it does not
    settle for this report.}
    \label{tab:stenosis-model-literature-comparison}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{@{}p{0.17\textwidth}p{0.12\textwidth}p{0.13\textwidth}p{0.18\textwidth}p{0.15\textwidth}p{0.16\textwidth}p{0.18\textwidth}@{}}
        \toprule
        \textbf{Source / class}
            & \textbf{State variables}
            & \textbf{Wall law}
            & \textbf{Stenosis treatment}
            & \textbf{Numerical method}
            & \textbf{Target}
            & \textbf{Limitation for this report} \\
        \midrule
        Classical stenosis-loss formulas
        \cite{YoungTsai1973SteadyStenosis,YoungTsai1973UnsteadyStenosis}
            & Flow and pressure-loss variables.
            & Usually prescribed geometry.
            & Empirical steady or unsteady loss terms.
            & Reduced algebraic/dynamic closure.
            & Pressure-drop prediction.
            & Not a distributed area--flow solver or velocity-profile comparison. \\
        Formaggia / Quarteroni 1D area--flow models
        \cite{FormaggiaEtAl2003OneDimensionalBloodFlow,QuarteroniFormaggia2004CardiovascularModeling}
            & Section area, flow, and pressure.
            & Pressure--area relation or coupled wall model.
            & Geometry enters through reference area and sources.
            & 1D hyperbolic finite-element/finite-volume formulations.
            & Arterial wave and network dynamics.
            & Provides the modeling framework, not this stenosis-specific implementation. \\
        Boileau benchmark solvers
        \cite{BoileauEtAl2015Benchmark}
            & 1D pressure, area, and flow.
            & Benchmark-specific arterial wall laws.
            & Network and vessel benchmark cases.
            & Cross-code numerical benchmark.
            & Solver agreement on prescribed cases.
            & Benchmark completion does not verify the selected C23/C40 stenosis realization. \\
        Canic extended 1D stenosis model
        \cite{CanicEtAl2024Extended1DStenosis}
            & Radius-squared area coordinate and flow.
            & Elastic Canic--Koiter-type pressure law.
            & Variable-radius distributed terms.
            & Reduced distributed 1D model.
            & Stenotic reduced-model dynamics.
            & The present solver is an $R_{\max}$-normalized Canic-derived variant, not a literal reproduction. \\
        OpenBF-style arterial workflow
        \cite{BenemeritoEtAl2024OpenBF}
            & 1D network pressure/flow states.
            & Declared vessel and terminal closures.
            & Network/vessel benchmark realization.
            & Open implementation and benchmark tooling.
            & Reproducible arterial simulations.
            & Useful for workflow comparison, not a matched resolved 3D stenosis reference. \\
        1D--3D pressure-ratio comparisons
        \cite{Blanco2018FFR1D3D,LuccaEtAl2025FFRPrediction,ShaikhEtAl2025CTFFR}
            & Pressure/flow observables and reduced indices.
            & Study-specific wall and boundary assumptions.
            & Coronary stenosis or CT-derived geometry.
            & Cross-dimensional or surrogate comparison.
            & FFR or CT-FFR-style quantities.
            & Shows the need for matched observables; this report compares velocity only. \\
        \bottomrule
    \end{tabular}%
    }
\end{table}

\subsection{Numerical Verification of 1D Arterial Solvers}
\label{subsec:verification-literature}

Verification asks whether the implementation solves the declared equations and
emits reproducible numerical records. For hyperbolic finite-volume solvers,
relevant evidence includes conservation diagnostics, CFL reporting,
positivity/limiting records, grid and timestep refinement, boundary-regime
checks, exact equilibria, and manufactured solutions
\cite{LeVeque2002FiniteVolumeMethodsForHyperbolicProblems}. Arterial solver
benchmarks provide a useful cross-code reference for wave propagation and
network behavior, but they do not replace problem-specific verification of a
selected stenosis realization
\cite{BoileauEtAl2015Benchmark,BenemeritoEtAl2024OpenBF}.

This distinction matters for the present report. A benchmark-stage completion
row, an output-consistency check, or a cross-integrator discrepancy can be
useful evidence about the computational realization, but it is weaker than a
manufactured-solution convergence table or a demonstrated discrete equilibrium.
A fixed-area characteristic boundary rule also requires evidence that the sign
condition under which it is applied holds in the sampled run. Verification
therefore has to persist diagnostic quantities, not only final plots.

\subsection{Cross-Dimensional Comparison and Validation}
\label{subsec:cross-dimensional-comparison-review}

A 1D--3D comparison is meaningful only after the compared observable has been
mapped to a common quantity. For velocity, this can mean reducing a resolved
3D velocity field to section averages, radial profiles, or point/slab samples.
For pressure and FFR-style quantities, the required maps are different and
must include proximal/distal pressure observations and time-averaging
conventions \cite{Blanco2018FFR1D3D,LuccaEtAl2025FFRPrediction,ShaikhEtAl2025CTFFR}.
Agreement under one operator is not agreement under another.

Validation is still a separate evidential category. A comparison with another
simulation can identify cross-model discrepancies, but it does not establish
truth unless the reference model and its boundary, wall, material, and history
data are themselves accepted for the target question. The present work
therefore uses cross-model velocity discrepancy language and reserves
\emph{error} for exact solutions, manufactured solutions, or accepted
reference data under matched conditions.

\subsection{Literature Synthesis and Research Gap}
\label{subsec:literature-synthesis-gap}

The reviewed literature supports a calibrated gap. Stenosis-aware reduced
models are useful because they are cheap, auditable, and compatible with
distributed network studies. Their outputs, however, depend on closure choices
and numerical realization. Published 1D, 3D, multiscale, and pressure-ratio
studies also show that verification, cross-dimensional comparison, and
physical validation answer different questions.

The present work addresses a practical reproducibility and observation-operator
gap for this implementation of an extended 1D stenosis model. It specifies the
solver-coordinate map, the Canic-derived model variant, targeted verification
evidence, and a declared plane--tetrahedron velocity comparison operator. A
stronger field-wide novelty or validation claim would require broader evidence
that comparable auditable implementations and matched 1D--3D operators do not
already exist, together with matched pressure or experimental data. That
broader claim is outside this report.
