\medskip
\noindent\textbf{Abstract.}\par\smallskip
This report reviews mathematical models and numerical methods used to simulate
blood flow, with emphasis on large-vessel and stenotic settings. The review
starts from continuum balance laws and incompressible Navier--Stokes models,
then compares three-dimensional CFD, fluid--structure interaction, axisymmetric
and two-dimensional reductions, distributed one-dimensional area--flow models,
lumped zero-dimensional networks, geometrical multiscale coupling, and emerging
learned surrogate approaches. The review is organized around source-to-source
contrasts: which equations and state variables are retained, which physical
effects are closed, which data are required, which numerical evidence is
reported, and which observables are native or reconstructed. Across those
tiers, the report separates the modeling choices that determine interpretation:
rheology, wall mechanics, geometry, inlet and outlet data, numerical
discretization, verification, validation, and observable definitions.

An idealized stenosis computation is used as a worked example of those review
principles. The case study uses the \path{StenosisHemodynamics} implementation
of an $R_{\max}$-normalized Canic-derived one-dimensional area--flow model, with
solver variables $a=R^2$ and $q=Q_{\mathrm{phys}}/\pi$. It records the model
contract, finite-volume realization, manufactured-solution verification record,
geometry-rest equilibrium test, production diagnostics, and a
plane--tetrahedron observation operator for comparing one-dimensional output
with available resolved three-dimensional velocity data. The MMS record over
$0\le t\le0.002$ s gives spatial $L^2$ rates near 1.70 for area and 1.50 for
flow, with accepted timestep and CFL reported separately. The zero-inlet rest
audit shows that the scheme is not exactly well balanced: final production-grid
rest defects are $0.0384$ and $0.0634\,\mathrm{cm}^3/\mathrm{s}$ for C23 and
C40, falling to $0.00973$ and $0.01605\,\mathrm{cm}^3/\mathrm{s}$ on the
$N=800$ audit grid. The exact-time $0.9995$ s 1D--3D section comparison gives
RMS mean-velocity discrepancies of 0.664 and 1.87 cm/s, maximum discrepancies
of 2.35 and 9.92 cm/s, and unresolved axial variation in extracted 3D flow of
7.7\% and 20.2\% of the case means. Radial-profile output is excluded pending
operator-test and reducer reconciliation. The case study illustrates how model
reduction, closure choices, equilibrium preservation, and matched observation
operators affect the meaning of computed blood-flow quantities.
