\section{Software and Reproducibility Record}
\label{app:code-and-ai-use}

This report is built from the \LaTeX{} sources rooted at
\path{final-report.tex}. The solver package is the root Julia package
\texttt{StenosisHemodynamics}, with package identity in \path{Project.toml},
resolved Julia dependencies in \path{Manifest.toml}, and the package entry point
at \path{src/StenosisHemodynamics.jl}. The command-line surface used by the
report is \path{scripts/stenosis-hemodynamics}, which dispatches through
\path{scripts/stenosis-hemodynamics.jl} and the root Julia project.

\paragraph*{Source-control and archive record.}
This supervised manuscript refresh was prepared on the local
\texttt{master} branch. The exact source tree for a submitted PDF is the git
commit containing this appendix and the refreshed report artifact. At the start
of the refresh, the previous committed state was:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
branch: master
previous_head_commit: 3b77c8b99ed3939dff6e4e6998095d7211338a22
previous_head_subject: Address executive assessment revisions
previous_head_commit_date: 2026-06-19T18:49:58-07:00
local_tag_available: masters-report-final-2026-06-19 -> 1a15a52d6e986f08d745c9e40ecbcd04d4475104
git_remote: none configured in the inspected checkout
public_repository_url: not declared
archival_doi: not declared
\end{lstlisting}
The local release manifest remains the concrete source-access record rather
than a final public archive record:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
reproducibility/release-manifest.json
\end{lstlisting}
No stable public repository URL, immutable archive URL, or archival DOI was
available from the local metadata in \path{CITATION.cff} or
\path{reproducibility/release-manifest.json}. Unless a later public release adds
one, access to \texttt{StenosisHemodynamics} is by the repository owner's
source-control policy.

\paragraph*{Computational environment.}
The checked Julia environment and local hardware record were:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
manifest_julia_version: 1.12.6
launcher_julia_version: julia version 1.12.6
operating_system: macOS 26.5.1, Darwin 25.5.0, arm64
hardware: Apple M1 Max, 10 logical CPUs, 32 GiB memory
\end{lstlisting}
The repository README states the intended portable execution surface as a macOS
or Linux shell with Julia 1.12 or newer, Python/Pipenv for auxiliary tooling,
and a TeX distribution with \texttt{latexmk} and \texttt{biber}.

\paragraph*{Review provenance.}
The narrative literature-hardening pass was completed on 19~June~2026 and
considered English-language records from 1973--2026. Representative targeted
queries were ``one-dimensional blood flow benchmark,'' ``well-balanced blood
flow finite volume,'' ``3D 1D fractional flow reserve comparison,''
``cardiovascular fluid structure interaction verification,'' and
``hemodynamics identifiability surrogate.'' The working evidence base consisted
of 68 BibLaTeX entries, 74 logged source records, a 132-row evidence matrix, and
a derived review-depth matrix. After triage, 38 source records were treated as
current cited evidence, 7 as report-adjacent scope checks, 22 as future-work
context, 2 as background only, and 5 as deferred pending metadata review. Blank
\texttt{bib\_key} records were not used for citation.

\paragraph*{Build command.}
The report build command is:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
python3 scripts/build_report.py --outdir /tmp/masters-report-build
\end{lstlisting}
Julia package tests are run through the repository launcher:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
./scripts/julia-release test/runtests.jl
\end{lstlisting}

\paragraph*{Comparison-asset regeneration command.}
When the optional raw 3D inputs are restored locally, the tracked comparison
assets are regenerated with:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
./scripts/stenosis-hemodynamics compare-3d \
  --data-root simulations/data/3d/canic_case3 \
  --output-dir simulations/output/3d_comparison/full_t1_native_nx400 \
  --nx 400 \
  --target-time 0.9995 \
  --time-atol 1e-3 \
  --overwrite \
  --publish-report-assets \
  --report-assets-dir figures/static/static/data/stenosis-comparison \
  --radial-bin-counts 10,20,40 \
  --radial-radius-modes current,reference
\end{lstlisting}
This command reproduces the currently tracked comparison assets; it accepts the
available $0.9995$ s 3D velocity field and runs the 1D comparison to the same
target time. The remaining timing limitation is transient-history matching, not
the final timestamp.

\paragraph*{Optional operator and grid-sensitivity evidence.}
The synthetic cross-section operator validation table is regenerated without
raw 3D inputs:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
./scripts/stenosis-hemodynamics operator-validation \
  --output-dir figures/static/static/tables/stenosis-comparison \
  --summary-csv figures/static/static/data/stenosis-comparison/cross-section-operator-validation.csv \
  --summary-tex figures/static/static/tables/stenosis-comparison/cross_section_operator_validation.tex \
  --overwrite
\end{lstlisting}
This run uses an in-memory tetrahedron with constant and affine axial fields.
It verifies the synthetic plane-cut reduction used by the cross-section
quadrature workflow and is independent of the optional HDF5/XDMF velocity
fields.

When the optional raw 3D inputs are available, the publication-scope grid
sensitivity is:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
./scripts/stenosis-hemodynamics compare-3d \
  --data-root simulations/data/3d/canic_case3 \
  --output-dir simulations/output/3d_comparison/grid_sensitivity_C23_C40 \
  --target-time 0.9995 \
  --time-atol 1e-3 \
  --nxs 200,400,800 \
  --section-count 200 \
  --radial-bins 20 \
  --no-svg \
  --overwrite \
  --grid-summary-csv figures/static/static/data/stenosis-comparison/grid-sensitivity-summary.csv \
  --grid-summary-tex figures/static/static/tables/stenosis-comparison/grid_sensitivity_summary.tex
\end{lstlisting}
The manuscript treats this as an output-sensitivity record for the retained
C23/C40 section means and physical flows. It does not change the matched-data
boundary: wall, boundary, material, current-geometry, and transient-history
metadata for the resolved cases remain unmatched or unpersisted.

Package-benchmark assets are regenerated separately from the PDF:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
./scripts/stenosis-hemodynamics benchmark \
  --profile overnight \
  --output-dir simulations/output/package_benchmark/overnight-YYYYMMDD \
  --overwrite \
  --include-resolved3d \
  --publish-report-assets

pipenv run python scripts/render_package_benchmark_figures.py \
  --benchmark-dir simulations/output/package_benchmark/overnight-YYYYMMDD
\end{lstlisting}
Those rows are secondary implementation-health records, not the principal
comparison method.

\paragraph*{Raw 3D inputs and manifest.}
The raw upstream XDMF/HDF5 velocity inputs used for
Section~\ref{sec:resolved-3d-comparison} are not archived in this report
package. The expected local input root is:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
simulations/data/3d/canic_case3/
\end{lstlisting}
When available, that directory is restored from this upstream repository,
commit, and subtree:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
https://github.com/qcutexu/Extended-1D-AQ-system
056a9da2b36b480691f18025d242d2c00f6e7180
case3_all_3d_results/
\end{lstlisting}
The report archives generated CSV and PGFPlots-ready static assets, not the raw
fields. Current/deformed 3D geometry, displacement application before plane
cuts, wall/boundary/material histories, and earlier transient histories remain
unresolved unless a later release record adds those data.

The tracked release manifest is stored outside the thesis PDF:
\begin{lstlisting}[basicstyle=\ttfamily\footnotesize,breaklines=true]
reproducibility/release-manifest.json
\end{lstlisting}
That manifest records the build command, raw-input policy, upstream 3D input
pointer, comparison-asset regeneration command, and the absence of a declared
public repository URL or archival DOI. No populated standalone SHA-256
file-hash manifest was present in the tracked source tree inspected for this
revision; per-run workflow manifests record SHA-256 output hashes when those
workflows are regenerated.
