\section{Numerical Methods Details}
\label{app:numerical-methods-details}
This appendix supplies the numerical-method detail behind
Section~\ref{subsubsec:cross-sectional-reduction-selected-1d-model} and the
comparison boundary in
Section~\ref{subsec:model-hierarchy-comparison-boundary}. It is supporting
method detail, not a claim of external-solver parity for the implemented
solver.
The continuous identities are read under
Convention~\ref{conv:foundation-standing-regularity}; computed quantities are
read through the following numerical-record convention.

\begin{convention}[Computed realization record]
\label{conv:computed-realization-record}
A computed realization is interpreted only through the model, closure,
discretization, grid, timestep policy, boundary-state rule, output operator,
and diagnostic metadata recorded for that realization. Symbols such as
$D_t$, $D_z$, numerical fluxes, residual arrays, and sampled norms denote
recorded discrete objects unless a section explicitly switches back to a
continuum identity.
\end{convention}

The time interval, output rectangle, and norm notation follow
Definition~\ref{def:foundation-time-and-compact-set-notation} and
Definition~\ref{def:foundation-norm-conventions}; discrete residuals use the
recorded $D_t$, $D_z$, grid, and quadrature weights rather than unnamed
continuum derivatives.

\subsection{Fixed-wall stationary-Stokes refinement record}
\label{app:num-stationary-stokes-refinement}

The local three-dimensional refinement study is limited to a fixed-wall
stationary-Stokes problem on the smooth idealized stenosis geometry. The public
entrypoint \path{run_stationary_stokes_refinement} constructs generated Gridap
tetrahedral meshes from the same $C^\infty$ radius profile used by the reduced
model, solves a pressure-drop-driven stationary Stokes problem, evaluates
section-averaged velocity and pressure from the finite-element field, and
compares those averages with the resistance/projection quantities used by the
stationary-Stokes initializer. This keeps the initializer behavior unchanged:
the study is an additional diagnostic record, not a replacement for the 1D
initial-condition contract.

For a Newtonian fixed-wall Stokes field, the sampled stress and wall-traction
diagnostics use
\begin{flalign*}
    \quad
    \boldsymbol\sigma(\vu,p)
    &=
    -p\mathbf I+\mu\left(\nabla\vu+\nabla\vu^\top\right), &&
    \mu=\rho\nu,\\
    \quad
    \mathbf t_w
    &=
    \boldsymbol\sigma\mathbf n, &
    \boldsymbol\tau_w
    &=
    \mathbf t_w-(\mathbf t_w\cdot\mathbf n)\mathbf n. &&
\end{flalign*}
Here $\mathbf n$ is the analytic outward normal of the $C^\infty$ wall profile.
The reported means and maxima are fixed-wall fluid-stress diagnostics sampled
from the finite-element solution. They are not structural displacement, wall
velocity, structural stress, or resolved FSI interface-work outputs.

\subsection{Elastic-potential balance-law form}
\label{app:num-elastic-potential-form}

The reduced state is $\mathbf U=(A,Q)^\top$. For a member of the
$\mathcal W_{\mathrm{elastic1D}}$ balance-law contract with prescribed
momentum-flux correction $\alpha$, the reference elastic-potential form is
\begin{flalign*}
    \quad
    \pd_t A+\pd_z Q &=0, &&\\
    \quad
    \pd_t Q
    +
    \pd_z \left(\alpha\frac{Q^2}{A}+\Psi(A,z,t)\right)
    &=
    S_{\mathrm{geom}}(A,z,t)
    -
    C_f(z,A,Q)\frac{Q}{A}. &&
\end{flalign*}
The elastic potential satisfies
\begin{flalign*}
    \quad
    \pd_A \Psi(A,z,t)
    =
    \frac{A}{\rho} \pd_A p_{\mathrm g}(A,z,t), &&
\end{flalign*}
and the geometry source is
\begin{flalign*}
    \quad
    S_{\mathrm{geom}}(A,z,t)
    =
    \left.\pd_z \Psi(A,z,t)\right|_A
    -
    \frac{A}{\rho}
    \left.\pd_z p_{\mathrm g}(A,z,t)\right|_A. &&
\end{flalign*}
For the selected Canic--Koiter wall-law member, assuming
$p_{\mathrm{ext}}$ is independent of $A$,
\begin{flalign*}
    \quad
    c^2(A,z)
    =
    \frac{A}{\rho} \pd_A p_{\mathrm g}(A,z,t)
    =
    \frac{\beta(z)\sqrt A}{2\rho A_0(z)}. &&
\end{flalign*}
Under the selected solver-coordinate map $A=\pi a$ and
$A_0=\pi R_0^2$, the implemented member corresponds to
$\beta(z)=\sqrt{\pi}K R_0(z)^2/R_{\max}^2$. Therefore
\begin{flalign*}
    \quad
    c^2(a,z)
    =
    \frac{K}{2\rho R_{\max}^2}\sqrt a
    &&
\end{flalign*}
for the evolution wall law used in the reported finite-volume runs.
Freeze $z$, the wall-closure data, and the source terms, and consider the
homogeneous $(A,Q)$ flux Jacobian associated with
$(Q,\alpha Q^2/A+\Psi(A,z,t))^\top$. On a positive-area state $A>0$, with
flow rate $Q$, mean velocity $Q/A$, prescribed momentum-flux correction
$\alpha$, and
$c^2(A,z)\defined(A/\rho)\pd_A p_{\mathrm g}(A,z,t)$, its eigenvalues are
\begin{flalign*}
    \quad
    \lambda_\pm(A,Q;z)
    =
    \alpha\frac{Q}{A}
    \pm
    \sqrt{c^2(A,z)+\alpha(\alpha-1)\left(\frac{Q}{A}\right)^2}. &&
\end{flalign*}
For the displayed elastic wall law, $c^2>0$ when $\beta>0$, $A>0$,
$A_0>0$, and $\rho>0$. Strict hyperbolicity of the homogeneous elastic
subsystem further requires the characteristic radicand
\begin{flalign*}
\quad
    c^2(A,z)+\alpha(\alpha-1)\left(\frac{Q}{A}\right)^2
    >0 &&
\end{flalign*}
to be positive on the reported state set. These speeds reduce to
$Q/A\pm c(A,z)$ in the special $\alpha=1$ reference case. The subcritical
case used in Assumption~\ref{ass:1d-subcritical-boundary-regime} is
$\lambda_-<0<\lambda_+$, so one characteristic enters through the inlet, one
characteristic enters through the outlet, and one scalar boundary condition is
imposed at each end of the vessel.
The displayed derivatives of $A_0$, $\beta$, $p_{\mathrm g}$, and
$\Psi$ are continuum notation for the model contract. A computed record may
realize them through explicitly recorded finite-difference or source-split
operators; changing that realization changes the computed map.

\subsection{Semi-discrete solver operator}
\label{app:num-current-solver-operator}

The implemented solver stores a radius-squared area coordinate in the variable
\texttt{A}. In this subsection write that coordinate as $a=R^2$. The
physical circular cross-sectional area is
$A_{\mathrm{phys}}=\pi a$. This is a solver-specific convention; the continuum
and classical 1D model sections keep $A$ for physical cross-sectional area.

For $N$ finite-volume cells, let
\begin{flalign*}
    \quad
    z_i=\left(i-\frac{1}{2}\right)\Delta z,\qquad
    \Delta z=\frac{L}{N},\qquad
    u_h(t)=
    \left(a_1(t),\ldots,a_N(t),q_1(t),\ldots,q_N(t)\right)^\top. &&
\end{flalign*}
The semi-discrete system exposed by \path{semidiscretize} and \path{rhs!} is
\begin{flalign*}
    \quad
    \dot u_h=\mathcal L_h(u_h,t;\theta), &&
\end{flalign*}
where $\theta$ denotes the recorded physical, geometry, grid, and solver
parameters. Componentwise,
\begin{flalign*}
    \quad
    \dot a_i
    &=
    -
    \frac{\widehat F^a_{i+1/2}-\widehat F^a_{i-1/2}}{\Delta z}, &&\\
    \quad
    \dot q_i
    &=
    -
    \frac{\widehat F^q_{i+1/2}-\widehat F^q_{i-1/2}}{\Delta z}
    +\widehat S_i. &&
\end{flalign*}
All divisions by $a$ use the positivity-regularized value of the
radius-squared area coordinate.

With $K=Eh/(1-\sigma^2)$, Canic conservative-form reference radius
$R_0^*=R_{\max}$, velocity-profile momentum factor
$\alpha_{\mathcal P}$, and
$\alpha_c(R_0')=-2(R_0')^2/35$, define
$\iota_{\mathcal M}=1$ for the \texttt{canic-extended-1d} forward-model
descriptor and $\iota_{\mathcal M}=0$ for
\texttt{classical-1d-no-slip}, the manifest token for the classical
parabolic-profile baseline. The semi-discrete full-state flux is
\begin{flalign*}
    \quad
    \mathbf F(a,q,z)
    =
    \left(
        q,\,
        \left(\alpha_{\mathcal P}+\iota_{\mathcal M}\alpha_c(R_0'(z))\right)\frac{q^2}{a}
        +\frac{K}{3\rho R_{\max}^2}a^{3/2}
    \right)^\top. &&
\end{flalign*}
At each interface,
\begin{flalign*}
    \quad
    \widehat{\mathbf F}_{i+1/2}
    =
    \frac{1}{2}\left[
        \mathbf F(\mathbf U^-_{i+1/2},z_{i+1/2})
        +
        \mathbf F(\mathbf U^+_{i+1/2},z_{i+1/2})
    \right]
    -
    \frac{1}{2}s_{i+1/2}
    \left(\mathbf U^+_{i+1/2}-\mathbf U^-_{i+1/2}\right), &&
\end{flalign*}
where $\mathbf U=(a,q)^\top$. The MUSCL finite-volume method limits the conservative
solver variables componentwise. For an interior cell,
\begin{flalign*}
    \quad
    \delta_i^a
    &=
    \operatorname{minmod}(a_i-a_{i-1},\,a_{i+1}-a_i),
    &
    \delta_i^q
    &=
    \operatorname{minmod}(q_i-q_{i-1},\,q_{i+1}-q_i), &&\\
    \quad
    \mathbf U^-_{i+1/2}
    &=
    (a_i+\tfrac12\delta_i^a,\,
     q_i+\tfrac12\delta_i^q)^\top,
    &
    \mathbf U^+_{i+1/2}
    &=
    (a_{i+1}-\tfrac12\delta_{i+1}^a,\,
     q_{i+1}-\tfrac12\delta_{i+1}^q)^\top. &&
\end{flalign*}
The active limiter is the two-argument minmod limiter, equivalent to the
$\theta=1$ member of the generalized minmod family. Boundary-adjacent cell
slopes are set to zero and paired with the ghost states described in
Appendix~\ref{app:num-boundary-state-realization}. Positivity limiting is
applied to reconstructed $a$ states before flux evaluation. The cell source is
not MUSCL-reconstructed; it is evaluated at cell centers using the recorded
$D_z^h$ differences below. The first-order finite-volume method is the separate
path that passes neighboring cell and ghost states directly to the same
Rusanov flux. The local numerical speed is
\begin{flalign*}
    \quad
    s_{i+1/2}
    =
    \max\left\{
        s_h(\mathbf U^-_{i+1/2},z_{i+1/2}),
        s_h(\mathbf U^+_{i+1/2},z_{i+1/2})
    \right\}, &&
\end{flalign*}
with
\begin{flalign*}
    \quad
    s_h(a,q,z)
    &=
    \max\left\{
        \left|\alpha_{\mathrm{eff}}u-\chi\right|,
        \left|\alpha_{\mathrm{eff}}u+\chi\right|
    \right\}, &&\\
    \quad
    \alpha_{\mathrm{eff}}
    &=
        \alpha_{\mathcal P}+\iota_{\mathcal M}\alpha_c(R_0'(z)),\qquad
    u=\frac{q}{a}, &&\\
    \quad
    \chi^2
    &=
    \left(\alpha_{\mathrm{eff}}u\right)^2
    -\alpha_{\mathrm{eff}}u^2
    +\frac{K}{2\rho R_{\max}^2}\sqrt a. &&
\end{flalign*}
This standard Rusanov/MUSCL operator is not presented here as an exactly
well-balanced discretization of the geometry-rest state. In particular, at
$q_i=0$ with spatially varying $a_i=R_0(z_i)^2$, the numerical viscosity in
the mass flux generally contributes a nonzero discrete residual unless an
equilibrium reconstruction or path-conservative flux is added. Thus
``source-balanced'' in the main text refers to the continuous flux/source pair;
this report does not claim
$\mathcal L_h(R_0(z_i)^2,0)=0$ to machine precision for the displayed
operator.

The source term uses the specified centered interior and one-sided boundary
difference $D_z^h$ on the cell averages. Let $g_{\mathcal P}$ denote the
profile shear factor,
\begin{flalign*}
    \quad
    \dot\gamma_{1D,i}=g_{\mathcal P}\frac{|q_i|}{a_i\sqrt{a_i}},
    \qquad
    \nu_{\mathrm{eff},i}^{\mathcal R}
    =
    \frac{
        \eta_{\mathrm{eff}}^{\mathcal R}(\dot\gamma_{1D,i})
    }{\rho},
    &&
\end{flalign*}
and
$\alpha_c'(z)=-4R_0'(z)R_0''(z)/35$. The implemented source is
\begin{flalign*}
    \quad
    \widehat S_i
    &=
    -2\nu_{\mathrm{eff},i}^{\mathcal R}g_{\mathcal P}\frac{q_i}{a_i}
    +
    \frac{K}{\rho R_{\max}^2}a_iR_0'(z_i)
    -
    \iota_{\mathcal M}a_i\,\widehat{\pd_z p_2}_i
    +
    \iota_{\mathcal M}\frac{q_i^2}{a_i}\alpha_c'(z_i), &&\\
    \quad
    \widehat{\pd_z p_2}_i
    &=
    \nu_{\mathrm{eff},i}^{\mathcal R} g_{\mathcal P}
    \left[
        \frac{q_iR_0''(z_i)}{a_iR_0(z_i)}
        -
        \frac{q_i(R_0'(z_i))^2}{a_iR_0(z_i)^2}
        +
        \frac{(D_z^h q)_iR_0'(z_i)}{a_iR_0(z_i)}
        -
        \frac{q_i(D_z^h a)_iR_0'(z_i)}{a_i^2R_0(z_i)}
    \right]. &&
\end{flalign*}
The $\widehat{\pd_z p_2}_i$ term is the density-scaled discrete derivative of
the Canic variable-radius viscous pressure correction. It is tied to the
\texttt{canic-extended-1d} realization and is identically inactive in the
model-family specification when $\iota_{\mathcal M}=0$, while the elastic
stiffness terms above come from the selected Canic--Koiter wall law.
For the \texttt{newtonian} descriptor, this reduces to the constant value
$\nu_{\mathrm{eff},i}^{\mathcal R}=\nu$. For the non-Newtonian closures,
the effective viscosity is evaluated with the shear-rate regularization and any
declared interval projection before division by density, then held locally
fixed in the analytic $p_2$ derivative. The realized generalized-Newtonian
$p_2$ contribution is therefore a computational closure rather than a fully
rederived variable-viscosity pressure correction.
For the power-family profile,
$g_{\mathcal P}=\gamma+2=\alpha_{\mathcal P}/(\alpha_{\mathcal P}-1)$.
The flat profile is not evaluated through that singular limit; it is assigned
$\alpha_{\mathcal P}=1$ and an explicit finite $g_{\mathcal P}$.

\subsection{Time-integration realization}
\label{app:num-time-integration-realization}

The native time integrator advances the semi-discrete operator with the declared
third-order strong-stability-preserving Runge--Kutta map
\begin{flalign*}
    \quad
    u_h^{n+1}
    =
    \Phi_{\Delta t_n}^{\mathrm{SSPRK3}}
    \left(u_h^n;\mathcal L_h\right). &&
\end{flalign*}
In expanded form, using $\Pi$ for the recorded positive-area projection on
the $a$-components,
\begin{flalign*}
    \quad
    u^{(1)} &= \Pi\left(u^n+\Delta t_n\mathcal L_h(u^n,t^n)\right), &&\\
    \quad
    u^{(2)} &= \Pi\left(
        \frac{3}{4}u^n
        +\frac{1}{4}\left[
            u^{(1)}+\Delta t_n\mathcal L_h(u^{(1)},t^n+\Delta t_n)
        \right]\right), &&\\
    \quad
    u^{n+1} &= \Pi\left(
        \frac{1}{3}u^n
        +\frac{2}{3}\left[
            u^{(2)}+\Delta t_n\mathcal L_h(u^{(2)},t^n+\Delta t_n)
        \right]\right). &&
\end{flalign*}
The native step size is
\begin{flalign*}
    \quad
    \Delta t_n
    =
    \min\left\{
        \Delta t_{\max},
        \frac{\mathrm{CFL}\,\Delta z}{\max_i s_h(a_i^n,q_i^n,z_i)},
        T_{\mathrm{final}}-t^n
    \right\}. &&
\end{flalign*}
The SciML realization constructs an \texttt{ODEProblem} for the same
$\dot u_h=\mathcal L_h(u_h,t;\theta)$ and applies the selected
OrdinaryDiffEq policy recorded in the solve specification. Thus comparisons
between these realizations change the time integrator, not the semi-discrete
spatial operator. The native fully discrete path applies the stage projection
$\Pi$, whereas the SciML path integrates the unprojected ODE system. If
$\Pi$ activates, the compared paths are no longer identical fully discrete
problems; the archived comparison records do not persist projection-activation
counts.

\subsection{Boundary-state realization}
\label{app:num-boundary-state-realization}

The finite-volume boundary realization imposes one incoming boundary condition
at each boundary in the subcritical regime using an $\alpha=1$, fixed-area
characteristic approximation. This is the implemented boundary-state rule for
the variable-$\alpha_{\mathrm{eff}}$ solver, not an exact invariant derivation
for the full variable-radius balance law. Let
\begin{flalign*}
    \quad
    c_0=\left(\frac{K}{2\rho (R_0^*)^2}\right)^{1/2}
    =
    \left(\frac{K}{2\rho R_{\max}^2}\right)^{1/2}. &&
\end{flalign*}
At the inlet, the incoming flow is
$q_{\mathrm{in}}=R_0(0)^2\bar u_{\mathrm{in}}$. The outgoing invariant from
the first interior cell is
\begin{flalign*}
    \quad
    w_-
    =
    \frac{q_1}{a_1}-4c_0a_1^{1/4}, &&
\end{flalign*}
and the ghost coordinate $a_{\mathrm{in}}$ solves
\begin{flalign*}
    \quad
    \frac{q_{\mathrm{in}}}{a_{\mathrm{in}}}
    -4c_0a_{\mathrm{in}}^{1/4}
    =
    w_-. &&
\end{flalign*}
At the outlet, the radius-squared coordinate is fixed to
$a_{\mathrm{out}}=R_0(L)^2$. The outgoing invariant from the last interior
cell is
\begin{flalign*}
    \quad
    w_+
    =
    \frac{q_N}{a_N}+4c_0a_N^{1/4}, &&
\end{flalign*}
and the ghost flow is
\begin{flalign*}
    \quad
    q_{\mathrm{out}}
    =
    a_{\mathrm{out}}w_+
    -4c_0a_{\mathrm{out}}^{5/4}. &&
\end{flalign*}
The boundary update therefore constrains the incoming information and leaves
the outgoing information tied to the interior state. Boundary-quality plots and
reflected-wave sidecars are diagnostics until explicit thresholds are admitted.

\subsection{Residual diagnostics}
\label{app:num-residual-diagnostics}

Stored residuals are post-processing diagnostics on sampled fields, not proof
of the continuous PDE residual. The symbols $D_t$ and $D_z$ in this
subsection denote the recorded temporal and spatial finite-difference operators
on the saved grid, not weak derivatives. For a stored field sequence, the
semi-discrete defect is read as
\begin{flalign*}
    \quad
    \mathcal E_h^n
    =
    D_t u_h^n-\mathcal L_h(u_h^n,t^n;\theta). &&
\end{flalign*}
Componentwise, this gives
\begin{flalign*}
    \quad
    \mathcal E^a_i(t^n)
    =
    D_t a_i^n-\mathcal L^a_{h,i}(u_h^n,t^n;\theta),
    \qquad
    \mathcal E^q_i(t^n)
    =
    D_t q_i^n-\mathcal L^q_{h,i}(u_h^n,t^n;\theta). &&
\end{flalign*}
The E1 diagnostic summary reports these residuals on reviewed interior cells
with recorded finite-field, positivity, CFL, and pressure-drop sign checks. The
numbers support the dataset contract and local consistency reading; they do not
establish a convergence theorem or external parity. Any residual norm or error
norm reported from these arrays is a sampled norm tied to the numerical record,
rather than an unqualified continuum $L^p(K)$ norm.
