\section{Introduction and Report Scope}
\label{sec:intro}

Coronary and arterial stenoses motivate a basic computational question: how do
geometry, wall properties, rheology, and boundary data determine the state of
blood flow through an idealized narrowed vessel? Clinical pressure-ratio
measurements motivate the question because anatomical narrowing alone does not
determine functional obstruction. This report, however, is a mathematical and
numerical study, not a clinical validation study.
\todo{ADVISOR[MAJOR]: Narrow the opening question to the actual deliverable.
The current wording suggests a parametric study over geometry, wall properties,
rheology, and boundary data, while the numerical evidence is two Newtonian
\(T=1\) velocity comparisons.}
\red{This report asks which continuum, closure, boundary-data, and
output-operator assumptions are needed to interpret a first controlled \(T=1\)
comparison between a selected 1D stenosis model and resolved 3D velocity data.}

The report fixes the vocabulary needed to move from continuum assumptions, to
the three-dimensional incompressible Navier--Stokes and generalized-Newtonian
settings, to reduced 0D/1D/2D hemodynamic model classes, and finally to a
selected one-dimensional numerical study of an idealized stenosis. Its
contribution is twofold: it organizes the relevant mathematical and numerical
literature, and it frames a controlled computation for a smooth idealized
stenosed vessel.

Pressure-ratio quantities in this report are computed model outputs, not
clinical FFR or CT-FFR measurements, unless an external measurement comparison
is explicitly introduced. Convention~\ref{conv:appendix-pressure-reference-ratio}
in Appendix~\ref{app:domain-specific-notation} records the pressure-reference,
gauge-pressure, high-flow-window, and admissibility conventions used for any
reduced pressure-ratio output.



\subsection*{Literature Review Roadmap}

\addcontentsline{toc}{subsection}{Literature Review Roadmap}

\begin{enumerate}
    \item Section~\ref{subsec:continuum-constitutive-assumptions} fixes the
    continuum, kinematic, balance-law, and constitutive vocabulary used later in
    the report.
    \item Section~\ref{subsec:governing-equations-model-hierarchy} identifies
    the governing equations, boundary-data vocabulary, and model hierarchy.
    \item Later literature-review sections use this vocabulary to compare
    classical finite-difference, finite-volume, and finite-element approaches
    for hemodynamic models.
    \item The numerical-study portion then applies the selected reduced model
    to a controlled idealized stenosis computation.
\end{enumerate}
\todo[inline]{ADVISOR[BLOCKER]: The roadmap promises later literature-review
sections comparing classical numerical methods, but \texttt{final-report.tex}
only includes the introduction, the \(T=1\) comparison, and appendices. Either
add those sections or revise this roadmap to the actual report contents before
submission.}
\red{The remaining main-text section uses this vocabulary to interpret a first
\(T=1\) 1D--3D velocity-output comparison for two idealized stenosis cases.}

\input{sections/01-intro/blood-continuum}
\input{sections/01-intro/governing-equations}
\todo{ADVISOR[MAJOR]: The clinical pressure--flow motivation currently appears
after the continuum and model hierarchy even though it motivates the problem.
Move this subsection before the technical subsections, or add a bridge
explaining why the reader must first read the mathematical setup.}
\red{The clinical motivation is stated after the mathematical setup because the
report first fixes which pressure quantities are model outputs rather than
measurements.}
\input{sections/01-intro/pressure-flow-motivation}
