\section{Closure Choices, Data, and Hemodynamic Observables}
\label{sec:modeling-closures}
\label{sec:modeling-closures-observables}

Choosing a model tier does not close a hemodynamic problem. The mathematical
state must be paired with constitutive, wall, geometric, boundary, data, and
observable conventions. These choices determine which terms appear in the
equations and which reported quantities are native states, derived fields,
diagnostics, or cross-tier observation maps. They are therefore part of the
mathematical model, not secondary implementation detail.

\subsection{Rheology and Effective-Viscosity Modeling}
\label{subsec:constitutive-wall-closures}

Blood is often modeled as an incompressible Newtonian fluid in large-vessel
simulations, especially when the target quantities are dominated by moderate to
high shear rates and the uncertainty in geometry or boundary data is already
large. Non-Newtonian categories remain important in stenotic, small-vessel, or
low-shear regimes because viscosity may depend on shear rate, hematocrit, and
diameter. Pries, Neuhaus, and Gaehtgens make the point from tube-flow evidence:
apparent blood viscosity is not a single universal constant once vessel scale
and hematocrit enter the measurement
\cite{PriesEtAl1992BloodViscosityTubeFlow}. Carreau and Carreau--Yasuda laws
therefore do more than fill a taxonomy slot; they encode shear-thinning with
limiting viscosities. Casson laws introduce yield-stress-like behavior, and
power-law models express scale-free shear dependence. Computational studies of
Carreau--Yasuda and intracranial stenosis rheology then show the consequence
for reporting: changing the viscosity law can alter the hemodynamic quantity
being interpreted, so the selected rheology must be carried with the observable
\cite{KumarEtAl2019CarreauYasudaHemodynamics,LiuEtAl2021ICASRheology}.

\begin{definition}[Effective-viscosity closure catalog]
\label{def:rheology-closure-catalog}
A reduced realization that selects a rheology family $\mathcal R$ must state
the scalar effective dynamic-viscosity law
$\eta_{\mathrm{eff}}^{\mathcal R}(\dot\gamma)$ before its friction, pressure,
or shear diagnostics can be interpreted. At the review level, the closure
contract consists of a dynamic-viscosity law, the shear-rate proxy on which it
is evaluated, the conversion $\nu_{\mathrm{eff}}^{\mathcal R}
=\eta_{\mathrm{eff}}^{\mathcal R}/\rho$, and any regularization or bounds used
near zero shear. The case-study implementation names these rheology families as
\begin{flalign*}
    \quad
    \texttt{newtonian},\quad
    \texttt{carreau},\quad
    \texttt{carreau-yasuda},\quad
    \texttt{casson},\quad
    \texttt{power-law}. &&
\end{flalign*}
Appendix~\ref{app:num-rheology-descriptor-details} gives the case-study
parameterization and regularization formulas. The listed families are
closure options used to make reduced rheology choices explicit, not
universal blood constants.
\end{definition}

\subsection{Wall Laws and Structural Assumptions}
\label{subsec:wall-laws-structural-assumptions}

Wall modeling is a separate closure axis. A fixed-wall model prescribes the
lumen boundary and solves only the fluid problem. A reduced compliant-wall
model uses an algebraic or differential pressure--area relation to close a 1D
or network state. A resolved FSI model solves structural wall motion and
couples it to the fluid at the interface
\cite{FormaggiaEtAl2003OneDimensionalBloodFlow,FormaggiaEtAl2007FSICoupling}.
The source contrast matters. Formaggia, Lamponi, and Quarteroni use a wall law
to close reduced pressure--area dynamics, while Formaggia, Moura, and Nobile
treat wall motion as an interface-coupled FSI state. Clinical imaging work on
coronary distensibility reinforces the same distinction from the observable
side: wall compliance can be a measured or estimated quantity rather than a
hidden numerical constant \cite{KelleEtAl2011CoronaryDistensibility}. The same
rheology can be paired with any of these wall choices, so rheology and wall
mechanics should not be merged into a single modeling label.

Membrane terminology belongs to the structural side of this split. A Newtonian
or generalized-Newtonian law describes the fluid stress, whereas an isotropic
elastic wall law describes how the vessel wall resists deformation. In this
report, a membrane model is a thin-wall structural closure: through-thickness
stress and bending are collapsed into a surface or centerline relation in which
fluid traction or transmural pressure drives normal or radial displacement and
elastic stiffness resists it. The quasi-static Canic-style relation used later,
$C_0\eta_r=f_r$, is this reduced radial balance in wall-law variables. It can
provide a wall-motion scale or structural closure, but it is not by itself a
full transient moving-boundary FSI model.

A wall-law statement must also say which configuration and pressure convention
it uses. Reference area or radius belongs to the unloaded, imaged, or otherwise
declared baseline geometry; current area or radius belongs to the deformed
state being solved or reconstructed. Pressure may be absolute, gauge, or
transmural, and a transmural law must state the external pressure convention.
Thickness, elastic modulus, prestress or residual stress, and viscoelastic
memory change the meaning of the same pressure--area curve. Reduced wall laws
therefore differ from resolved structural states: the former close pressure,
area, and wave speed inside a 1D model, while the latter solve wall motion and
stress. Wall parameters may also compensate for geometry or boundary
uncertainty, so they are not automatically identifiable from pressure and flow
alone.

\subsection{Stenosis Geometry and Reference-State Choices}
\label{subsec:geometry-boundary-observable-closures}

A stenosis can enter the mathematics as a resolved lumen narrowing, an
axisymmetric radius profile, a 1D reference-area variation, an empirical
pressure-loss relation, or a network resistance. Distributed 1D formulations
make the reference area and its derivatives part of the balance-law closure,
while resolved models represent the same narrowing through the computational
domain. The extended 1D stenosis formulation is important here because it
argues that variable-radius effects can alter the reduced balance law itself;
they are not captured by simply inserting a narrowed healthy-radius profile
into an otherwise standard area--flow model
\cite{CanicEtAl2024Extended1DStenosis}. Geometry is therefore not only a mesh
or image artifact: it determines whether throat effects appear through domain
shape, source terms, pressure loss, or lumped parameters.

Classical idealized stenosis studies occupy the corresponding evidence role.
Young and Tsai's steady and unsteady model-stenosis experiments connect
controlled narrowing to pressure-loss and flow-structure effects under declared
geometry and flow conditions
\cite{YoungTsai1973SteadyStenosis,YoungTsai1973UnsteadyStenosis}. They are
therefore useful evidence for obstruction mechanisms and pressure-loss framing,
not patient-specific stenosis validation or clinical severity inference.
Severity must also be defined by radius, diameter, or area rather than treated
as interchangeable. A 23\% radius reduction, diameter reduction, and area
reduction imply different lumen changes. Patient-specific geometry adds another
layer: Gambaruto and coauthors found, in one cerebral aneurysm case, that
reconstructions differing within imaging resolution produced sensitivity
greater than but comparable to rheology choice and WSS differences of 3--18\%
\cite{GambarutoEtAl2011GeometryRheologySensitivity}. That result is
case-specific, but it supports the broader reporting rule: geometry
reconstruction is part of the model uncertainty. Idealized geometry is useful
for mechanism isolation; patient-specific geometry is required for
patient-specific inference.

\subsection{Inlet, Outlet, and Initial Data}
\label{subsec:inlet-outlet-initial-data}

Boundary data are model dependent. Inlets may be prescribed as velocity
profiles, mean-flow waveforms, pressure histories, or characteristic incoming
data. Outlets may be prescribed as pressure, traction, resistance, Windkessel,
impedance, reflection, or characteristic conditions. In multiscale models, the
same words also name interface contracts between tiers, not only external
boundary conditions
\cite{QuarteroniVeneziani2003GeometricalMultiscale,ShiEtAl2011BloodFlowModelReview}.
Quarteroni and Veneziani treat such data as coupling variables in a
geometrical multiscale system, while Shi, Lawford, and Hose emphasize how
network and terminal choices close lumped or distributed pressure--flow states.
For a single-vessel problem, the boundary is therefore not just a geometric
edge. A resolved 3D problem may prescribe velocity, flow, traction, pressure,
or wall-coupling information on inlet, outlet, and wall portions, while a 1D
model applies incoming information to area--flow or characteristic variables.
The case study later states its boundary rule as part of its model contract,
because boundary data close the mathematical problem.
For subcritical 1D flow, the conceptual characteristic count is one incoming
piece of information at each end of a single vessel. Prescribed pressure,
prescribed flow, traction, impedance/Windkessel, characteristic, and multiscale
coupling data therefore impose different mathematical information. Boundary
conditions can encode omitted vascular dynamics, and initial transients must be
distinguished from a periodic or statistically steady operating assumption.

\subsection{Pressure, Flow, Velocity, WSS, and FFR Observables}
\label{subsec:hemodynamic-observables}

Observables should be named at the level where they are actually defined.
Pressure may be an absolute pressure, gauge pressure, cross-sectional mean
pressure, diagnostic pressure reconstructed from a wall law, or terminal
compartment pressure. Flow may be a surface integral, a 1D state variable, or a
network edge variable. Velocity may be a resolved vector field, a section
average, an assumed radial profile, or a sampled centerline quantity. WSS is a
traction-derived wall quantity requiring stress and wall geometry, not a
generic name for any reduced shear proxy
\cite{JohnEtAl2017WallShearStressVectorForm}. Displacement belongs to a wall
or structural state, so it is native to resolved FSI or a structural model and
only indirectly represented by a reduced pressure--area wall law.
Thus $p(t,\vx)$, $\vu(t,\vx)$, and stress or shear components denote continuum
field values only in the continuum tier. Reported scalars, curves, table
entries, and visualizations are observables produced by maps from a field or
reduced state unless a pointwise field evaluation is explicitly declared.

Clinical pressure-flow quantities add another observable layer. FFR literature
separates anatomical stenosis severity from functional severity. The FAME trial
source supplies the clinical contrast between angiographic appearance and
functional assessment, while FFR overview literature defines the diagnostic as
a pressure-ratio quantity rather than a generic flow-model output
\cite{ToninoEtAl2010FAMEAngioFunctional,TebaldiEtAl2015FFROverview}.
CT-derived FFR sources then shift the same diagnostic into an image-derived
computational pipeline, so the observable depends on segmentation, boundary
assumptions, solver model, and pressure-ratio extraction
\cite{NorgaardEtAl2014NXTFFRCT,ShaikhEtAl2025CTFFR}. The 1D--3D FFR comparison
literature makes the cross-tier version of the point: a pressure-ratio scalar
can be compared across model classes only after both states have been observed
through compatible pressure and flow conventions \cite{Blanco2018FFR1D3D}.

The operator forms used in the report are intentionally explicit observation
maps:
\begin{flalign*}
    \quad
    Q=\int_S \vu\cdot\nhat_S\,dA,\qquad
    \bar u=Q/A,\qquad
    \bar p=A^{-1}\int_S p\,dA, &&
\end{flalign*}
where $S$ is the declared section and $\bar p$ is a section representative, not
a pointwise pressure value. A pressure drop is
$\Delta p=p_{\mathrm{up}}-p_{\mathrm{down}}$, and a pressure ratio is
$p_{\mathrm{down}}/p_{\mathrm{up}}$ only after the pressure convention,
locations, and pressure observation maps are declared. A WSS vector is the
tangential projection of $\vT\nhat$ on the wall, with scalar indices such as
AWSS or OSI derived only after the vector convention and wall-output operator
are fixed. An FFR-style pressure ratio should not be called FFR unless the
physiological, boundary, and diagnostic assumptions of the FFR setting are part
of the model contract.

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Closure-contract axes that should remain explicit before comparing
    stenotic-hemodynamics models.}
    \label{tab:modeling-closures-observables}
    \begin{tabularx}{\textwidth}{@{}
        >{\raggedright\arraybackslash}p{0.16\textwidth}
        >{\raggedright\arraybackslash}p{0.23\textwidth}
        >{\raggedright\arraybackslash}p{0.23\textwidth}
        >{\raggedright\arraybackslash}X@{}}
        \toprule
        \textbf{Closure}
            & \textbf{Mathematical location}
            & \textbf{Required declaration}
            & \textbf{What this does not establish} \\
        \midrule
        Rheology
            & Stress law or 1D friction source.
            & $\eta$ or $\eta_{\mathrm{eff}}(\dot\gamma)$ and proxy.
            & A viscosity law alone does not determine pressure-ratio or WSS
            sensitivity; the observable and regime must be stated
            \cite{LiuEtAl2021ICASRheology}. \\
        Velocity profile
            & Momentum flux and wall friction.
            & Profile shape, $\alpha$, shear factor.
            & A profile closure does not produce a native resolved radial field
            or wall traction. \\
        Wall mechanics
            & Pressure--area law or structural PDE.
            & Reference/current geometry, stiffness, external pressure.
            & A wall law does not identify stiffness, prestress, or material
            memory unless those data are independently constrained. \\
        Geometry/reference state
            & Domain, reference radius, source terms, loss model.
            & Segmentation or analytic radius and severity definition.
            & A shared nominal severity does not remove reconstruction or
            reference-state uncertainty
            \cite{GambarutoEtAl2011GeometryRheologySensitivity}. \\
        Inlet data
            & Boundary state or interface.
            & Flow, pressure, velocity profile, characteristic data.
            & A prescribed inlet does not model omitted upstream dynamics unless
            it is supplied by a coupling model. \\
        Outlet data
            & Boundary state, terminal ODE, or impedance.
            & Pressure, traction, resistance, Windkessel, characteristic data.
            & A terminal label does not determine wave reflection or distal
            circulation without its parameters and assumptions. \\
        Junction law
            & Network interface.
            & Conservation and pressure/energy condition.
            & A junction condition does not make variables compatible across
            tiers unless pressure, flow, and energy conventions match. \\
        \bottomrule
    \end{tabularx}
\end{table}

\begin{table}[!htb]
    \centering
    \scriptsize
    \caption{Reporting conventions for reduced and cross-tier observables.}
    \label{tab:modeling-observable-reporting-conventions}
    \begin{tabularx}{\textwidth}{@{}
        >{\raggedright\arraybackslash}p{0.18\textwidth}
        >{\raggedright\arraybackslash}p{0.25\textwidth}
        >{\raggedright\arraybackslash}p{0.24\textwidth}
        >{\raggedright\arraybackslash}X@{}}
        \toprule
        \textbf{Observable convention}
            & \textbf{Required declaration}
            & \textbf{Affected output}
            & \textbf{What this does not establish} \\
        \midrule
        Pressure convention
            & Gauge/absolute/transmural reference.
            & Pressure drop and ratio.
            & A pressure convention does not define external pressure or
            upstream/downstream sites by itself. \\
        Flow or velocity map
            & Section, coordinate mode, quadrature, and averaging rule.
            & Flow, mean velocity, and discrepancy summaries.
            & A section mean does not establish resolved secondary-flow or WSS
            agreement. \\
        WSS convention
            & Stress, wall normal, projection.
            & WSS, AWSS, OSI.
            & A scalar WSS index does not identify the vector convention unless
            the traction projection is declared
            \cite{JohnEtAl2017WallShearStressVectorForm}. \\
        FFR-style observation
            & Upstream/downstream pressures and physiological contract.
            & FFR or FFR-style result.
            & A pressure ratio should not be called FFR without the physiological
            and boundary contract \cite{Blanco2018FFR1D3D}. \\
        \bottomrule
    \end{tabularx}
\end{table}

\begin{stenosisillustration}
\label{ex:stenosisillustration-modeling-closures}
The case study later in the report uses this closure framework in a bounded
way: Newtonian viscosity for the principal comparison; a parabolic velocity
profile for the main area--flow closure; an elastic pressure--area wall law; a
smooth analytic stenosis reference radius; a prescribed reference-area inlet
flow; and a fixed-area characteristic outlet approximation. The native states
are reduced area and flow, with mean velocity and pressure obtained through
declared coordinate and wall-law maps. The resolved-data comparison therefore
reads as a velocity-observation comparison, not as a direct WSS or
wall-displacement comparison.
\end{stenosisillustration}
